---
title: "Podocyte Meeting Poster Figures"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}

options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```

# Loading Datasets

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_all <- readRDS(here("outputs", "Alports_041525.rds"))

```

# The UMAP of glom cells

## Option 1

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DimPlot(SO_all) + 
  ylab("UMAP 2") +
  xlab("UMAP 1") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.line = element_line(size = 1, colour = "black"),
    axis.text = element_blank(),          # Remove axis text
    axis.ticks = element_blank(),         # Optional: Remove axis ticks
    text = element_text(size = 20)
  )



```

## Option 2

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width = 8}

SO_all@meta.data <- SO_all@meta.data %>% 
  mutate(class_poster_name = dplyr::case_when(
    class_merge == "EC"  ~ "Endothelial Cell",
    class_merge == "IMM"  ~ "Immune Cell",
    class_merge == "MES"  ~ "Mesangial Cell",
    class_merge == "PODO"  ~ "Podocyte",
    class_merge == "PEC"  ~ "Parietal Epithelial Cell",
    class_merge == "VSMC"  ~ "Vascular Smooth Muscle Cell"
    ))

SO_all@meta.data$class_poster_name <- factor(SO_all@meta.data$class_poster_name, levels = c("Endothelial Cell", "Immune Cell", "Mesangial Cell", "Podocyte", "Parietal Epithelial Cell", "Vascular Smooth Muscle Cell"))

DimPlot(SO_all, group.by = "class_poster_name") + 
  ylab("UMAP 2") +
  xlab("UMAP 1") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.line = element_line(size = 1, colour = "black"),
    axis.text = element_blank(),          # Remove axis text
    axis.ticks = element_blank(),         # Optional: Remove axis ticks
    text = element_text(size = 20)
  ) + 
  ggtitle("")

```

## Extra

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=6, fig.width=6}

markers.to.plot1 <- c(
                      "Flt1",        # Endo
                      "Ptprc",        # Immune
                      "Pdgfrb",       # Mesangial
                      "Nphs2",        # Podo        
                      "Ncam1",        # PEC
                      "Acta2"        # VSMC
)    

DotPlot(SO_all,
        features = markers.to.plot1,
        dot.scale = 10,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5,
        group.by = "class_poster_name") +
  coord_flip() +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1)) +  # <- this line tilts the x-axis labels
  xlab(NULL) +
  ylab(NULL)


```

# Cell Type DEG

a.	could you please reduce the width of the graph, I have only limited space? <br>
b.	I suggest removing "Alport" and "4M" on the x-axis and displaying only the cell type names. <br>
c.	I suggest renaming the graph title to "Glomerular Cell type DEGs". I can define Alport and 4M in the legend. <br> 


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=4, fig.height=4}

rm(list = ls())

load(here("outputs", "Alports_4M_Condition_DEGs.Rdata"))

# List the objects in the environment
  objects_in_environment <- ls()

data_frames <- Filter(is.data.frame, mget(ls()))

# Initialize an empty data frame to store the results
output_table <- data.frame(Data_Frame = character(),
                           Number_of_Obs = numeric(),
                           Number_of_Positive_Log2FC = numeric(),
                           Number_of_Negative_Log2FC = numeric(),
                           stringsAsFactors = FALSE)

# Loop through each data frame
for (df_name in names(data_frames)) {
  df <- data_frames[[df_name]]
  
  # Get the number of observations in the data frame
  num_obs <- nrow(df)
  
  # Count the number of observations with positive log2FC
  positive_count <- sum(df$avg_log2FC > 0)
  
  # Count the number of observations with negative log2FC
  negative_count <- sum(df$avg_log2FC < 0)
  
  # Create a row in the output table
  output_row <- data.frame(Data_Frame = df_name,
                           Number_of_Obs = num_obs,
                           Number_of_Positive_Log2FC = positive_count,
                           Number_of_Negative_Log2FC = negative_count,
                           stringsAsFactors = FALSE)
  
  # Append the row to the output table
  output_table <- rbind(output_table, output_row)
}

# Print the output table
print(output_table)

# Measure the total number of DEGs (sum of output_table$Number_of_Obs)
v1 <-sum(output_table$Number_of_Obs)
print("Total Number of DEGs")

print(v1)

output_table2 <- output_table %>% 
                 mutate(Data_Frame = sub(".*\\.", "", Data_Frame)) %>%
                 arrange(desc(Number_of_Obs)) %>%
                 mutate(Data_Frame = factor(Data_Frame, levels = unique(Data_Frame)))

output_table2$Data_Frame <- gsub("Alports_4M_", "", output_table2$Data_Frame)

p1 <- ggplot(output_table2, aes(x = Data_Frame)) +
  geom_bar(aes(y = Number_of_Positive_Log2FC), stat = "identity", fill = "coral2", width = 0.5) +
  geom_bar(aes(y = -Number_of_Negative_Log2FC), stat = "identity", fill = "dodgerblue2", width = 0.5) +
  labs(x = "", y = "Number of DEGs") +
  ggtitle("Glomerular Cell Type DEGs") +
  scale_y_continuous(labels = function(x) abs(x)) +
  annotate("text", x = Inf, y = Inf, label = "Positive Log2FC", color = "coral2",
           hjust = 1, vjust = 1, size = 8) +
  annotate("text", x = Inf, y = Inf, label = "Negative Log2FC", color = "dodgerblue2",
           hjust = 1, vjust = 2.1, size = 8) +
  theme_minimal() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 20),
        axis.text = element_text(color = "black", size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title = element_text(color = "black", size = 16))

p1

```

## Option 2

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

 p1 <- ggplot(output_table2, aes(x = Data_Frame)) +
  geom_bar(aes(y = Number_of_Positive_Log2FC), stat = "identity", fill = "coral2") +
  geom_bar(aes(y = -Number_of_Negative_Log2FC), stat = "identity", fill = "dodgerblue2") +
  labs(x = "", y = "Number of DEGs") +
  ggtitle("Glomerular Cell Type DEGs") +
  scale_y_continuous(labels = function(x) abs(x)) +
  annotate("text", x = Inf, y = Inf, label = "Positive Log2FC", color = "coral2",
           hjust = 1, vjust = 1, size = 8) +
  annotate("text", x = Inf, y = Inf, label = "Negative Log2FC", color = "dodgerblue2",
           hjust = 1, vjust = 2.1, size = 8) +
  theme_minimal() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 24),
        axis.text = element_text(color = "black", size = 14,),
        axis.text.x = element_text(color = "black", size = 14, angle = 45, hjust = 1, vjust = 1),
        axis.title = element_text(color = "black", size = 20,))

p1

```

# GEC UMAP with subclusters

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}

SO_ec <- readRDS(here("outputs", "Alports_Glom_EC.rds"))

DimPlot(SO_ec) + 
  ylab("UMAP 2") +
  xlab("UMAP 1") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.line = element_line(size = 1, colour = "black"),
    axis.text = element_blank(),          # Remove axis text
    axis.ticks = element_blank(),         # Optional: Remove axis ticks
    text = element_text(size = 20)
  ) + 
  ggtitle("")


```

# GEC subclusters

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}


SO_ec@meta.data <- SO_ec@meta.data %>% 
  mutate(class_poster_group_name = dplyr::case_when(
    group == "WT"  ~ "WT",
    group == "4M_Alp"  ~ "4M Alports",
    group == "4M_Alp_1EV"  ~ "4M Alports 1EV",
    group == "4M_Alp_2EV"  ~ "4M Alports 2EV",
    group == "6M_Alp"  ~ "6M Alports"
    ))

t1 <- table(SO_ec@meta.data$seurat_clusters, SO_ec@meta.data$class_poster_group_name)[, c("WT", "4M Alports", "4M Alports 1EV", "4M Alports 2EV", "6M Alports")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic() +
  theme(axis.line = element_line(size = 1, colour = "black"),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        text = element_text(size=20),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(fill = "Cell Type") +
  xlab(NULL)


```


# GEC UMAP by group ( 6M Alport removed) 

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=5}

SO_ec_2 <- subset(SO_ec, subset = group != "6M_Alp")

DimPlot(SO_ec_2, reduction = "umap", split.by = "group", group.by = "seurat_clusters") +
  xlab("UMAP 1") +
  ylab("UMAP 2") +
  ggtitle("") +
  theme_classic() +
  theme(
    axis.line = element_line(size = 1, colour = "black"),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_blank(),  # <- This removes facet titles
    text = element_text(size = 20)
  )

```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

