---
title: "Alports Merging and Filtering"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}

options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```

# Loading Filtered Seurat Object
These files are post-DoubletFinder.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO1 <- readRDS(here("outputs", "Alports_WT.rds"))
SO2 <- readRDS(here("outputs", "Alports_4M.rds"))
SO3 <- readRDS(here("outputs", "Alports_4M_EV.rds"))
SO4 <- readRDS(here("outputs", "Alports_4M_2EV.rds"))
SO5 <- readRDS(here("outputs", "Alports_6M.rds"))

```

# Nuclei Entering the Analysis Pipeline
The following shows the number of nuclei post-QC and post-DoubletFinder. Numbers in `# (10X Count)` are the starting numbers for reference.  
Values appear in the Environment when running the `rmd`.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO1.nuc <- nrow(SO1@meta.data) 
SO1.nuc
SO2.nuc <- nrow(SO2@meta.data) 
SO2.nuc
SO3.nuc <- nrow(SO3@meta.data) 
SO3.nuc
SO4.nuc <- nrow(SO4@meta.data) 
SO4.nuc
SO5.nuc <- nrow(SO5@meta.data) 
SO5.nuc


head(SO1@meta.data)
```

# Merging Replicates

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- merge(SO1, y = c(SO2, SO3, SO4, SO5), project = "Alports")

```

# Add Metadata

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

meta <- read.xlsx(here("datasets", "metadata.xlsx"), sheet = 2)

meta2 <- meta  %>% select(c("Sample.Name", "batch", "group", "group_num", "alports", "ev")) 

df <- FetchData(SO, "sample") %>% rownames_to_column(var = "CellID")

df <- left_join(df, meta2, by = c("sample" = "Sample.Name")) %>% column_to_rownames(var = "CellID")

SO <- AddMetaData(SO, df)

head(SO@meta.data)

SO@meta.data$group <- factor(SO@meta.data$group, levels = c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp"))

```



## Nuclei Filtering and Batch Comparison

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=10}


VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "group",
        split.by = "sample")

VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "group",
        split.by = "sample", pt.size = 0)

```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}


VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "group")


VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "group", pt.size = 0)


SO

VlnPlot(SO, "percent.mt", group.by = "class_merge", split.by = "group", pt.size = 0) +
  ylim(0, 2)


SO <- readRDS(here("outputs", "Alports_041525.rds"))


s.genes <- Seurat::cc.genes.updated.2019$s.genes
g2m.genes <- Seurat::cc.genes.updated.2019$g2m.genes

convert_human_to_mouse <- function(gene_list) {
    output = c()
    mouse_human_genes = read.csv("https://www.informatics.jax.org/downloads/reports/HOM_MouseHumanSequence.rpt",sep="\t")

    for(gene in gene_list) {
          class_key = (mouse_human_genes %>% filter(Symbol == gene & Common.Organism.Name == "human"))[['DB.Class.Key']]
          if( !identical(class_key, integer(0)) ) {
            human_genes = (mouse_human_genes %>% filter(DB.Class.Key == class_key & Common.Organism.Name=="mouse, laboratory"))[,"Symbol"]
            for(human_gene in human_genes) {
                output = rbind(c(gene, human_gene), output)
            }
          }
     }
     return (output)
}

s.genes  <- convert_human_to_mouse(s.genes)
s.genes <- as.data.frame(s.genes)
s.genes <- s.genes$V2

g2m.genes  <- convert_human_to_mouse(g2m.genes)
g2m.genes <- as.data.frame(g2m.genes)
g2m.genes <- g2m.genes$V2

SO <- CellCycleScoring(
  object = SO,
  s.features = s.genes,
  g2m.features = g2m.genes,
  set.ident = TRUE  # optional: sets Seurat identity to cell cycle phase
)

table(SO$Phase, SO@meta.data$group)

DimPlot(SO, group.by = "Phase", label = TRUE) +
  labs(title = "Cell Cycle Phase")

SO2 <- subset(SO, subset = class_merge == "PODO")

DimPlot(SO2)

table(SO2$Phase, SO2@meta.data$group)

t1 <- table(SO2$Phase, SO2@meta.data$group)[, c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic()



SO3 <- subset(SO, subset = class_merge == "EC")

DimPlot(SO3)

table(SO3$Phase, SO3@meta.data$group)

t1 <- table(SO3$Phase, SO3@meta.data$group)[, c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic()



```


## Chart of Nuclei Number

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

meta <- SO@meta.data

# Keep only categorical columns (character or factor)
meta_categorical <- meta %>%
  select(sample, group, where(~ is.character(.x) || is.factor(.x)))

# Pivot and count
summary_df <- meta_categorical %>%
  pivot_longer(cols = -c(sample, group), names_to = "meta_col", values_to = "meta_value") %>%
  filter(meta_value == "Singlet") %>%
  group_by(sample, group) %>%
  summarise(n_cells = n(), .groups = "drop")

summary_df

summary_df$sample <- factor(summary_df$sample, 
                            levels = summary_df %>%
                              arrange(group, sample) %>%
                              pull(sample) %>%
                              unique())

ggplot(summary_df, aes(x = sample, y = n_cells, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal(base_size = 14) +
  labs(title = "Number of Cells per Sample",
       y = "Number of Cells",
       x = "Sample",
       fill = "Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_classic(base_size = 14)


# Pivot and count
summary_df2 <- meta_categorical %>%
  pivot_longer(cols = -c(group), names_to = "meta_col", values_to = "meta_value") %>%
  filter(meta_value == "Singlet") %>%
  group_by(group) %>%
  summarise(n_cells = n(), .groups = "drop")

summary_df2

ggplot(summary_df, aes(x = group, y = n_cells, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal(base_size = 14) +
  labs(title = "Number of Cells per Group",
       y = "Number of Cells",
       x = "Sample",
       fill = "Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme_classic()


```

## Standard Seurat Pipeline

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

SO <- SCTransform(SO) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = 2) %>%
    RunUMAP(dims = 1:30)

DimPlot(SO, reduction = "umap", label = TRUE)

DimPlot(SO, reduction = "umap", group.by = "group")

DimPlot(SO, reduction = "umap", group.by = "class_merge")


```

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width=14}

DimPlot(SO, reduction = "umap", split.by = "group")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

t1 <- table(SO@meta.data$class_merge, SO$group)[, c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic()

t1 <- table(SO@meta.data$class_merge, SO$sample)[, c("452", "453", "454", "365", "372", "474", "362", "368", "475", "364", "375", "473", "A", "B", "C")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic()

```

## Filtering Non-Glom clusters

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=12}

markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Cdh13",        # dTL1
                      "Slc14a2",      # dTL2
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs2",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b",        # Uro
                      "Top2a"         # Proliferation
)
                      
DotPlot(SO,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

DimPlot(SO, group.by = "class_merge")

FeaturePlot(SO, "Ehd3")
FeaturePlot(SO, "Tspan7")
FeaturePlot(SO, "Pi16")
FeaturePlot(SO, "Plvap")
FeaturePlot(SO, "Igfbp3", order = T)
FeaturePlot(SO, "Npr3", order = T)
FeaturePlot(SO, "Cryab", order = T)
FeaturePlot(SO, "Aqp1", order = T)

Idents(SO) <- SO$class_merge

markers.EC <- FindMarkers(SO, ident.1 = "4M_Alp", ident.2 = "WT" , group.by = "group", subset.ident = "EC", logfc.threshold = 0.25, min.pct = 0.1)

VlnPlot(SO, "Fabp4", group.by = "group", idents = "EC")
VlnPlot(SO, "Fkbp5", group.by = "group", idents = "EC")
VlnPlot(SO, "Zbtb16", group.by = "group", idents = "EC")
VlnPlot(SO, "Gda", group.by = "group", idents = "EC")
VlnPlot(SO, "Ly6a", group.by = "group", idents = "EC")
VlnPlot(SO, "Lpl", group.by = "group", idents = "EC")
VlnPlot(SO, "Insr", group.by = "group", idents = "EC")
VlnPlot(SO, "Itgb3", group.by = "group", idents = "EC")

VlnPlot(SO, "Fosb", group.by = "group", idents = "EC", split.by = "sample")

VlnPlot(SO, "Col4a5", group.by = "group", idents = "PODO", split.by = "sample")

VlnPlot(SO, "Col4a5", group.by = "group", idents = "PODO")

FeaturePlot(SO, "Col4a5", split.by = "group", order = T)

```

# Save Object

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(SO, here("outputs", "Alports_041525.rds"))

```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

