---
title: "Alports Merging and Filtering"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}

library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)



options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```
# Introduction

## Changes based on meeting with Sargis on 4/22/2025

1) Save Alports_Glom_EC.rds file <br>
2) Change arguments for FindMarkers to `logfc.threshold = 0.1` <br>
3) Write DEG lists to excel file <br>
4) Write Pathways list to excel file <br>
5) Add comparisons of additional groups (WT vs. 6M, 4M vs. 4M_1EV) <br>


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- readRDS(here("outputs", "Alports_041525.rds"))

DimPlot(SO)

VlnPlot(SO, "Col4a5", group.by = "group", idents = "PODO")

VlnPlot(SO, "Col4a4", group.by = "group", idents = "PODO")

VlnPlot(SO, "Col4a3", group.by = "group", idents = "PODO")

```

# Subset Endothelial Cells

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_EC <- subset(SO, idents = c("EC"))

DimPlot(SO_EC)

SO_EC@meta.data
```

## Recluster and Filter

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results='hide'}

SO_EC <- SCTransform(SO_EC) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = 2) %>%
    RunUMAP(dims = 1:30)

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DimPlot(SO_EC, label = T)

```

## Identifying EC subtypes

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10}

markers.to.plot.ec <- c("Pecam1",         # EC
                      "Flt1",      # EC
                      "Cdh5",      # EC
                      "Pi16",      # glom EC
                      "Ehd3",       # glom EC
                      "Cyp4b1",        # glom EC
                      "Tspan7",        # glom EC
                      "Cdh13",        #  glom EC
                      "Plvap",      # peri EC
                      "Igfbp3",      # peri EC
                      "Igf1",         # peri EC-m
                      "Cryab",        # peri EC-m
                      "Igfbp7",       # peri EC-m
                      "Cd36",       # peri EC-m
                      "Aqp1"        # peri EC-m

)


DotPlot(SO_EC,
features = markers.to.plot.ec,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```

## Checking non-EC's

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=10}


markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Cdh13",        # dTL1
                      "Slc14a2",      # dTL2
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs2",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b",        # Uro
                      "Top2a"         # Proliferation
)
                      
DotPlot(SO_EC,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}


SO_EC2 <- SO_EC

SO_EC2@meta.data <- SO_EC2@meta.data %>% 
  mutate(subclass_EC = dplyr::case_when(
    seurat_clusters == 0  ~ "EC_glom",
    seurat_clusters == 1  ~ "EC_glom",
    seurat_clusters == 2  ~ "EC_glom",
    seurat_clusters == 3  ~ "EC_glom",
    seurat_clusters == 4  ~ "EC_glom",
    seurat_clusters == 5  ~ "EC_glom",
    seurat_clusters == 6  ~ "EC_glom",
    seurat_clusters == 7  ~ "EC_glom",
    seurat_clusters == 8  ~ "EC_glom",
    seurat_clusters == 9  ~ "EC_glom",
    seurat_clusters == 10 ~ "EC_glom",
    seurat_clusters == 11 ~ "EC_glom",
    seurat_clusters == 12 ~ "EC_periM1",
    seurat_clusters == 13 ~ "EC_glom",
    seurat_clusters == 14 ~ "EC_periC1",
    seurat_clusters == 15 ~ "EC_glom",
    seurat_clusters == 16 ~ "EC_glom",
    seurat_clusters == 17 ~ "EC_glom",
    seurat_clusters == 18 ~ "EC_glom",
    seurat_clusters == 19 ~ "EC_glom",
    seurat_clusters == 20 ~ "EC_glom",
    seurat_clusters == 21 ~ "EC_glom",
    seurat_clusters == 22 ~ "EC_glom",
    seurat_clusters == 23 ~ "EC_glom",
    seurat_clusters == 24 ~ "EC_glom",
    seurat_clusters == 25 ~ "EC_glom",
    seurat_clusters == 26 ~ "EC_glom",
    seurat_clusters == 27 ~ "EC_periM2",
    seurat_clusters == 28 ~ "EC_periM3",
    seurat_clusters == 29 ~ "Doublet",
    seurat_clusters == 30 ~ "EC_glom",
    seurat_clusters == 31 ~ "IMM"
  ))

SO_EC2@meta.data$subclass_EC <- factor(SO_EC2@meta.data$subclass_EC , levels = c("EC_glom", "EC_periC1", "EC_periM1", "EC_periM2", "EC_periM3", "IMM", "Doublet"))

Idents(SO_EC2) <- SO_EC2@meta.data$subclass_EC

DimPlot(object = SO_EC2, reduction = "umap", group.by = "subclass_EC", label = TRUE)


DimPlot(object = SO_EC2, reduction = "umap", label = TRUE)

```


```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=10}

markers.to.plot.ec2 <- c("Pecam1",         # EC
                         "Ehd3",
                         "Plvap",      # peri EC
                         "Igfbp3",      # peri EC
                         "Cryab",        # peri EC-m
                         "Ptprc"        # Immune
                         
)


DotPlot(SO_EC2,
features = markers.to.plot.ec2,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```


## Recluster ECs

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

SO_EC_Clean <- subset(SO_EC2, subset = subclass_EC == "EC_glom")

DimPlot(SO_EC_Clean)

SO_EC_Clean <- SCTransform(SO_EC_Clean) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = .5) %>%
    RunUMAP(dims = 1:30)

DimPlot(SO_EC_Clean)

# saveRDS(SO_EC_Clean, file = here("outputs", "Alports_Glom_EC.rds"))

```

# DEGs

## 4M vs. WT

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

Idents(SO_EC_Clean) <- SO_EC_Clean@meta.data$subclass_EC

df <- FindMarkers(SO_EC_Clean, ident.1 = "4M_Alp", ident.2 = "WT", group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

dfx <- df %>% arrange(desc(avg_log2FC))

dfx

test <- head(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test, group.by = "group")

test2 <- tail(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test2, group.by = "group")


write.xlsx(dfx, file = here("outputs", "DEGs_Alports_Glom_EC_DEGs_4M_vs_WT.xlsx"), row.names = TRUE)

```

### Volcano Plot

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

df2 <- df %>% filter(p_val_adj < 0.05)

top5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = TRUE)[1:10]]
bottom5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = FALSE)[1:10]]

# Combine them into a single vector
selected_genes <- c(top5, bottom5)

EnhancedVolcano(df2,
                lab = rownames(df2),
                selectLab = selected_genes,
                title = (""),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 4, 
                legendIconSize = 4, 
                drawConnectors = T,
                max.overlap = 5,
                xlim = c(-2.5, 2.5),
                widthConnectors = 0.5) + 
  theme_classic(base_size = 16) +  # Increases overall font size
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.text = element_text(size = 14),  # Axis tick labels
        axis.title = element_text(size = 16, face = "bold"),  # Axis labels
        axis.line = element_line(size = 1.2),  # Increases axis border thickness
        legend.position = "none")

```

### Upregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

# rotate this plot around the y axis

dotplot(pos_go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_4M_vs_WT_upregulated.xlsx"))


```

### Downregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go) +
    ggtitle("") +
    theme_classic() + 
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "left",
        axis.text.y = element_text(hjust = 0, size = 10)) +
    scale_y_discrete(position = "right", 
                     labels = function(x) str_wrap(x, width = 25))  # Wrap y-axis labels to 2 lines

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_4M_vs_WT_downregulated.xlsx"))

```


##  4M_1xEV vs. 4M

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

Idents(SO_EC_Clean) <- SO_EC_Clean@meta.data$subclass_EC

df <- FindMarkers(SO_EC_Clean, ident.1 = "4M_Alp_1EV", ident.2 = "4M_Alp", group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

dfx <- df %>% arrange(desc(avg_log2FC))

dfx

test <- head(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test, group.by = "group")

test2 <- tail(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test2, group.by = "group")

write.xlsx(dfx, file = here("outputs", "DEGs_Alports_Glom_EC_DEGs_4M_1xEV_vs_4M.xlsx"), row.names = TRUE)


```

### Volcano Plot

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

df2 <- df %>% filter(p_val_adj < 0.05)

top5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = TRUE)[1:10]]
bottom5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = FALSE)[1:10]]

# Combine them into a single vector
selected_genes <- c(top5, bottom5)

EnhancedVolcano(df2,
                lab = rownames(df2),
                selectLab = selected_genes,
                title = (""),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 4, 
                legendIconSize = 4, 
                drawConnectors = T,
                max.overlap = 5,
                xlim = c(-2.5, 2.5),
                widthConnectors = 0.5) + 
  theme_classic(base_size = 16) +  # Increases overall font size
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.text = element_text(size = 14),  # Axis tick labels
        axis.title = element_text(size = 16, face = "bold"),  # Axis labels
        axis.line = element_line(size = 1.2),  # Increases axis border thickness
        legend.position = "none")
```

### Upregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

# rotate this plot around the y axis

dotplot(pos_go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_4M_1xEV_vs_4M_upregulated.xlsx"))



```

### Downregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go) +
    ggtitle("") +
    theme_classic() + 
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "left",
        axis.text.y = element_text(hjust = 0, size = 10)) +
    scale_y_discrete(position = "right", 
                     labels = function(x) str_wrap(x, width = 25))  # Wrap y-axis labels to 2 lines

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_4M_1xEV_vs_4M_downregulated.xlsx"))

```


##  4M_2xEV vs. 4M

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

Idents(SO_EC_Clean) <- SO_EC_Clean@meta.data$subclass_EC

df <- FindMarkers(SO_EC_Clean, ident.1 = "4M_Alp_2EV", ident.2 = "4M_Alp", group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

dfx <- df %>% arrange(desc(avg_log2FC))

dfx

test <- head(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test, group.by = "group")

test2 <- tail(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test2, group.by = "group")

write.xlsx(dfx, file = here("outputs", "DEGs_Alports_Glom_EC_DEGs_4M_2xEV_vs_4M.xlsx"), row.names = TRUE)


```

### Volcano Plot

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

df2 <- df %>% filter(p_val_adj < 0.05)

top5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = TRUE)[1:10]]
bottom5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = FALSE)[1:10]]

# Combine them into a single vector
selected_genes <- c(top5, bottom5)

EnhancedVolcano(df2,
                lab = rownames(df2),
                selectLab = selected_genes,
                title = (""),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 4, 
                legendIconSize = 4, 
                drawConnectors = T,
                max.overlap = 5,
                xlim = c(-2.5, 2.5),
                widthConnectors = 0.5) + 
  theme_classic(base_size = 16) +  # Increases overall font size
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.text = element_text(size = 14),  # Axis tick labels
        axis.title = element_text(size = 16, face = "bold"),  # Axis labels
        axis.line = element_line(size = 1.2),  # Increases axis border thickness
        legend.position = "none")
```

### Upregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

# rotate this plot around the y axis

dotplot(pos_go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_4M_2xEV_vs_4M_upregulated.xlsx"))



```

### Downregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go) +
    ggtitle("") +
    theme_classic() + 
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "left",
        axis.text.y = element_text(hjust = 0, size = 10)) +
    scale_y_discrete(position = "right", 
                     labels = function(x) str_wrap(x, width = 25))  # Wrap y-axis labels to 2 lines

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_4M_2xEV_vs_4M_downregulated.xlsx"))

```


## 6M vs. WT

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

Idents(SO_EC_Clean) <- SO_EC_Clean@meta.data$subclass_EC

df <- FindMarkers(SO_EC_Clean, ident.1 = "6M_Alp", ident.2 = "WT", group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

dfx <- df %>% arrange(desc(avg_log2FC))

dfx

test <- head(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test, group.by = "group")

test2 <- tail(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test2, group.by = "group")


write.xlsx(dfx, file = here("outputs", "DEGs_Alports_Glom_EC_DEGs_6M_vs_WT.xlsx"), row.names = TRUE)

```

### Volcano Plot

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

df2 <- df %>% filter(p_val_adj < 0.05)

top5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = TRUE)[1:10]]
bottom5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = FALSE)[1:10]]

# Combine them into a single vector
selected_genes <- c(top5, bottom5)

EnhancedVolcano(df2,
                lab = rownames(df2),
                selectLab = selected_genes,
                title = (""),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 4, 
                legendIconSize = 4, 
                drawConnectors = T,
                max.overlap = 5,
                xlim = c(-2.5, 2.5),
                widthConnectors = 0.5) + 
  theme_classic(base_size = 16) +  # Increases overall font size
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.text = element_text(size = 14),  # Axis tick labels
        axis.title = element_text(size = 16, face = "bold"),  # Axis labels
        axis.line = element_line(size = 1.2),  # Increases axis border thickness
        legend.position = "none")

```

### Upregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

# rotate this plot around the y axis

dotplot(pos_go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_6M_vs_WT_upregulated.xlsx"))


```

### Downregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go) +
    ggtitle("") +
    theme_classic() + 
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "left",
        axis.text.y = element_text(hjust = 0, size = 10)) +
    scale_y_discrete(position = "right", 
                     labels = function(x) str_wrap(x, width = 25))  # Wrap y-axis labels to 2 lines

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_6M_vs_WT_downregulated.xlsx"))

```



## 6M vs. 4M

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

Idents(SO_EC_Clean) <- SO_EC_Clean@meta.data$subclass_EC

df <- FindMarkers(SO_EC_Clean, ident.1 = "6M_Alp", ident.2 = "4M_Alp", group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

dfx <- df %>% arrange(desc(avg_log2FC))

dfx

test <- head(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test, group.by = "group")

test2 <- tail(rownames(dfx), n = 3)

VlnPlot(SO_EC_Clean, features = test2, group.by = "group")


write.xlsx(dfx, file = here("outputs", "DEGs_Alports_Glom_EC_DEGs_6M_vs_4M.xlsx"), row.names = TRUE)

```

### Volcano Plot

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

df2 <- df %>% filter(p_val_adj < 0.05)

top5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = TRUE)[1:10]]
bottom5 <- rownames(df2)[order(df2$avg_log2FC, decreasing = FALSE)[1:10]]

# Combine them into a single vector
selected_genes <- c(top5, bottom5)

EnhancedVolcano(df2,
                lab = rownames(df2),
                selectLab = selected_genes,
                title = (""),
                subtitle = NULL,
                caption = NULL,
                x = 'avg_log2FC', 
                legendLabels = NULL, 
                FCcutoff = 0.25, 
                y = 'p_val_adj', 
                labSize = 4, 
                legendIconSize = 4, 
                drawConnectors = T,
                max.overlap = 5,
                xlim = c(-2.5, 2.5),
                widthConnectors = 0.5) + 
  theme_classic(base_size = 16) +  # Increases overall font size
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        axis.text = element_text(size = 14),  # Axis tick labels
        axis.title = element_text(size = 16, face = "bold"),  # Axis labels
        axis.line = element_line(size = 1.2),  # Increases axis border thickness
        legend.position = "none")

```

### Upregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC < 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

# rotate this plot around the y axis

dotplot(pos_go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_6M_vs_WT_upregulated.xlsx"))


```

### Downregulated Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

DEG_list <- df2

markers <- DEG_list %>% rownames_to_column(var="SYMBOL")

head(markers, n = 50)

ENTREZ_list <- bitr(geneID = rownames(DEG_list),   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

markers <-  ENTREZ_list %>% inner_join(markers, by = "SYMBOL")

# Removing genes that are not statistically significant. 
markers <-  markers %>% dplyr::filter(p_val_adj < 0.05)
#head(markers, n = 50)

pos.markers <-  markers %>% dplyr::filter(avg_log2FC > 0) %>%  arrange(desc(abs(avg_log2FC)))
#head(pos.markers, n = 50)

pos.ranks <- pos.markers$ENTREZID[abs(pos.markers$avg_log2FC) > 0]
#head(pos.ranks)

pos_go <- enrichGO(gene = pos.ranks,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name

pos_go

dotplot(pos_go) +
    ggtitle("") +
    theme_classic() + 
    theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "left",
        axis.text.y = element_text(hjust = 0, size = 10)) +
    scale_y_discrete(position = "right", 
                     labels = function(x) str_wrap(x, width = 25))  # Wrap y-axis labels to 2 lines

df_pathways <- pos_go@result

write.xlsx(df_pathways, file = here("outputs", "Pathways_Alports_Glom_EC_DEGs_6M_vs_4M_downregulated.xlsx"))

```


# Identifying Reversal Genes

## Set the conditions

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Condition_Baseline <- "WT"

Condition_A <- "4M_Alp"
  
Condition_B <- "4M_Alp_2EV"

```

## Create DEG lists

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_A, ident.2 = Condition_Baseline, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df2 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_B, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df1

```

## Split by Up and Down

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1_pos <- df1 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df1_pos

df1_neg <- df1 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df1_neg

df2_pos <- df2 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df2_pos

df2_neg <- df2 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df2_neg


```

## Reversal Venn Diagrams

### Up -> Down

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_pos)
set2 <- rownames(df2_neg)
venn <- ggvenn(list(DEGA = set1, DEGB = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

df3 <- head(test$'A:B', 10)

for (i in df3) {
    f1 <- VlnPlot(SO_EC_Clean, i, group.by = "group")
    print(f1)
}

genelist <- test$'A:B'

genelist

df1_pos <- df1_pos %>% rownames_to_column(var = "gene")

# Filter using genelist
df1_pos_filtered <- df1_pos %>% 
  filter(gene %in% genelist)

df2_neg <- df2_neg %>% rownames_to_column(var = "gene")

# Filter using genelist
df2_neg_filtered <- df2_neg %>% 
  filter(gene %in% genelist)

df_filtered <- inner_join(df1_pos_filtered, df2_neg_filtered, by = "gene")

df_filtered$difference <- abs(df_filtered$avg_log2FC.x) - abs(df_filtered$avg_log2FC.y)

df_filtered <- df_filtered %>% arrange(abs(difference))

df_filtered  <- df_filtered  %>% column_to_rownames(var = "gene")

list2 <- head(rownames(df_filtered), 5)

for (i in list2) {
    f1 <- VlnPlot(SO_EC_Clean, i, group.by = "group")
    print(f1)
}


```

### Down -> Up

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_neg)
set2 <- rownames(df2_pos)
venn <- ggvenn(list(DEGA = set1, DEGB = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

df3 <- head(test$'A:B', 10)

for (i in df3) {
    f1 <- VlnPlot(SO_EC_Clean, i, group.by = "group")
    print(f1)
}

genelist <- test$'A:B'

genelist

df1_neg <- df1_neg %>% rownames_to_column(var = "gene")

# Filter using genelist
df1_pos_filtered <- df1_neg %>% 
  filter(gene %in% genelist)

df2_pos <- df2_pos %>% rownames_to_column(var = "gene")

# Filter using genelist
df2_neg_filtered <- df2_pos %>% 
  filter(gene %in% genelist)

df_filtered <- inner_join(df1_pos_filtered, df2_neg_filtered, by = "gene")

df_filtered$difference <- abs(df_filtered$avg_log2FC.x) - abs(df_filtered$avg_log2FC.y)

df_filtered <- df_filtered %>% arrange(abs(difference))

df_filtered  <- df_filtered  %>% column_to_rownames(var = "gene")

list2 <- head(rownames(df_filtered), 5)

for (i in list2) {
    f1 <- VlnPlot(SO_EC_Clean, i, group.by = "group")
    print(f1)
}



```


## Heatmap

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- AverageExpression(object = SO_EC_Clean, features = c(df3), group.by = 'group')$RNA

df

df <- as.data.frame(df)

# remove columns 4M_Alp_1EV and 6M_Alp

df <- df[, -c(3, 5)]

df <- t(scale(t(df)))

# convert df to tidy format

df_tidy <- df %>%
  as.data.frame() %>%
  rownames_to_column(var = "Gene") %>%
  pivot_longer(cols = -Gene, names_to = "group", values_to = "Expression")
               
# Graph with geom_tile

df_tidy$group <- factor(df_tidy$group, levels = c("WT", "4M_Alp", "4M_Alp_2EV"))

f4 <- ggplot(df_tidy, aes(x = group, y = Gene, fill = Expression)) +
  geom_tile(color = "black", size = 0.2, width = 0.97, height = 0.90) +  # Adds thin border and spacing between tiles
  scale_fill_distiller(palette = "RdYlBu") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    axis.ticks = element_blank(),         # Removes tick marks
    panel.grid = element_blank(),          # Removes grid lines
    legend.position = "bottom",            # Centers the legend below the plot
    legend.title = element_text(size = 14), # Increases legend title size
    legend.text = element_text(size = 12)   # Increases legend text size
  ) +
  labs(x = "Cell Type", y = "Gene", fill = "Relative Expression")

f4

```



## Heatmap 2

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- AverageExpression(object = SO_EC_Clean, features = c(df3), group.by = 'group')$RNA

df

df <- as.data.frame(df)

# remove columns 4M_Alp_1EV and 6M_Alp

df <- df[, -c(3, 5)]

df <- t(scale(t(df)))

# convert df to tidy format

df_tidy2 <- df %>%
  as.data.frame() %>%
  rownames_to_column(var = "Gene") %>%
  pivot_longer(cols = -Gene, names_to = "group", values_to = "Expression")
               
# Graph with geom_tile

df_tidy2$group <- factor(df_tidy2$group, levels = c("WT", "4M_Alp", "4M_Alp_2EV"))

f4 <- ggplot(df_tidy2, aes(x = group, y = Gene, fill = Expression)) +
  geom_tile(color = "black", size = 0.2, width = 0.97, height = 0.90) +  # Adds thin border and spacing between tiles
  scale_fill_distiller(palette = "RdYlBu") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    axis.ticks = element_blank(),         # Removes tick marks
    panel.grid = element_blank(),          # Removes grid lines
    legend.position = "bottom",            # Centers the legend below the plot
    legend.title = element_text(size = 14), # Increases legend title size
    legend.text = element_text(size = 12)   # Increases legend text size
  ) +
  labs(x = "Group", y = "Gene", fill = "Relative Expression")

f4

```


### Heatmap3

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df_tidy$set <- "A"
df_tidy2$set <- "B"

df_tidy3 <- rbind(df_tidy, df_tidy2)

df_tidy3$set <- factor(df_tidy3$set, levels = c("A", "B"))

df_tidy3 <- df_tidy3 %>%
  mutate(Gene = factor(Gene, levels = unique(Gene[order(set)])))

# Then plot
f4 <- ggplot(df_tidy3, aes(x = group, y = Gene, fill = Expression)) +
  geom_tile(color = "black", size = 0.2, width = 0.97, height = 0.90) +
  scale_fill_distiller(palette = "RdYlBu") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  labs(x = "Group", y = "Gene", fill = "Relative Expression")


f4


```

## Heatmap

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- AverageExpression(object = SO_EC_Clean, features = c(df3), group.by = 'group')$RNA

df

df <- as.data.frame(df)

# remove columns 4M_Alp_1EV and 6M_Alp

df <- df[, -c(5)]

df <- t(scale(t(df)))

# convert df to tidy format

df_tidy <- df %>%
  as.data.frame() %>%
  rownames_to_column(var = "Gene") %>%
  pivot_longer(cols = -Gene, names_to = "group", values_to = "Expression")
               
# Graph with geom_tile

df_tidy$group <- factor(df_tidy$group, levels = c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV"))

f4 <- ggplot(df_tidy, aes(x = group, y = Gene, fill = Expression)) +
  geom_tile(color = "black", size = 0.2, width = 0.97, height = 0.90) +  # Adds thin border and spacing between tiles
  scale_fill_distiller(palette = "RdYlBu") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    axis.ticks = element_blank(),         # Removes tick marks
    panel.grid = element_blank(),          # Removes grid lines
    legend.position = "bottom",            # Centers the legend below the plot
    legend.title = element_text(size = 14), # Increases legend title size
    legend.text = element_text(size = 12)   # Increases legend text size
  ) +
  labs(x = "Cell Type", y = "Gene", fill = "Relative Expression")

f4

```



## Heatmap 2

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df <- AverageExpression(object = SO_EC_Clean, features = c(df3), group.by = 'group')$RNA

df

df <- as.data.frame(df)

# remove columns 4M_Alp_1EV and 6M_Alp

df <- df[, -c(5)]

df <- t(scale(t(df)))

# convert df to tidy format

df_tidy2 <- df %>%
  as.data.frame() %>%
  rownames_to_column(var = "Gene") %>%
  pivot_longer(cols = -Gene, names_to = "group", values_to = "Expression")
               
# Graph with geom_tile

df_tidy2$group <- factor(df_tidy2$group, levels = c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV"))

f4 <- ggplot(df_tidy2, aes(x = group, y = Gene, fill = Expression)) +
  geom_tile(color = "black", size = 0.2, width = 0.97, height = 0.90) +  # Adds thin border and spacing between tiles
  scale_fill_distiller(palette = "RdYlBu") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    axis.ticks = element_blank(),         # Removes tick marks
    panel.grid = element_blank(),          # Removes grid lines
    legend.position = "bottom",            # Centers the legend below the plot
    legend.title = element_text(size = 14), # Increases legend title size
    legend.text = element_text(size = 12)   # Increases legend text size
  ) +
  labs(x = "Group", y = "Gene", fill = "Relative Expression")

f4

```


### Heatmap3

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df_tidy$set <- "A"
df_tidy2$set <- "B"

df_tidy3 <- rbind(df_tidy, df_tidy2)

df_tidy3$set <- factor(df_tidy3$set, levels = c("A", "B"))

df_tidy3 <- df_tidy3 %>%
  mutate(Gene = factor(Gene, levels = unique(Gene[order(set)])))

# Then plot
f4 <- ggplot(df_tidy3, aes(x = group, y = Gene, fill = Expression)) +
  geom_tile(color = "black", size = 0.2, width = 0.97, height = 0.90) +
  scale_fill_distiller(palette = "RdYlBu") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  labs(x = "Group", y = "Gene", fill = "Relative Expression")


f4


```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

