---
title: "Alports 4M Merging and Filtering"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}

options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```

# Loading Filtered Seurat Object
These files are post-DoubletFinder.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

One <- readRDS(here("outputs", "Alports_372.rds"))
Two <- readRDS(here("outputs", "Alports_474.rds"))
Three <- readRDS(here("outputs", "Alports_365.rds"))

```

# Nuclei Entering the Analysis Pipeline
The following shows the number of nuclei post-QC and post-DoubletFinder. Numbers in `# (10X Count)` are the starting numbers for reference.  
Values appear in the Environment when running the `rmd`.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
One.nuc <- nrow(One@meta.data) 
One.nuc
Two.nuc <- nrow(Two@meta.data) 
Two.nuc
Three.nuc <- nrow(Three@meta.data) 
Three.nuc

```

# Merging Replicates

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- merge(One, y = c(Two, Three), project = "Alports_4M")

```

## Nuclei Filtering and Batch Comparison

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}

head(SO@meta.data)

VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "sample")


VlnPlot(SO, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "sample", pt.size = 0)


SO

```


## Standard Seurat Pipeline

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.show='hide', fig.align = 'center'}

SO <- SCTransform(SO) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = 1) %>%
    RunUMAP(dims = 1:30)
```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DimPlot(SO, reduction = "umap", label = TRUE)

DimPlot(SO, reduction = "umap", group.by = "sample")


```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=12}

DimPlot(SO, reduction = "umap", split.by = "sample")

```

## Filtering Non-Glom clusters

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=12, fig.height=10}

markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Cdh13",        # dTL1
                      "Slc14a2",      # dTL2
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs2",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b",        # Uro
                      "Top2a"         # Proliferation
)
                      
DotPlot(SO,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```


## Annotate Clusters
### subclass

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}


SO2 <- SO

SO2@meta.data <- SO2@meta.data %>% 
  mutate(subclass_merge = dplyr::case_when(
    seurat_clusters == 0  ~ "EC_1",
    seurat_clusters == 1  ~ "DCT_1",
    seurat_clusters == 2  ~ "EC_2",
    seurat_clusters == 3  ~ "EC_3",
    seurat_clusters == 4  ~ "PT_1",
    seurat_clusters == 5  ~ "PT_2",
    seurat_clusters == 6  ~ "PODO",
    seurat_clusters == 7  ~ "EC_4",
    seurat_clusters == 8  ~ "IMM_1",
    seurat_clusters == 9  ~ "PT_3",
    seurat_clusters == 10 ~ "MES",
    seurat_clusters == 11 ~ "Doublet_1",
    seurat_clusters == 12 ~ "IMM_2",
    seurat_clusters == 13 ~ "DCT_2",
    seurat_clusters == 14 ~ "IMM_3",
    seurat_clusters == 15 ~ "PC",
    seurat_clusters == 16 ~ "IMM_4",
    seurat_clusters == 17 ~ "IC",
    seurat_clusters == 18 ~ "IMM_5",
    seurat_clusters == 19 ~ "PEC",
    seurat_clusters == 20 ~ "EC_5",
    seurat_clusters == 21 ~ "IMM_6",
    seurat_clusters == 22 ~ "TAL",
    seurat_clusters == 23 ~ "Doublet_2",
    ))

#SO2@meta.data$class.Merge <- factor(SO2@meta.data$class.Merge , levels = c("Fibroblasts", "Medullary Fibroblasts", "Proliferating", "Mesangial Cells", "Pericytes", "Efferent #Artery", "Afferent Artery", "JGA"))

DimPlot(object = SO2, reduction = "umap", group.by = "subclass_merge", label = TRUE)

```

### class


```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}



SO2@meta.data <- SO2@meta.data %>% 
  mutate(class_merge = dplyr::case_when(
    seurat_clusters == 0  ~ "EC",
    seurat_clusters == 1  ~ "DCT",
    seurat_clusters == 2  ~ "EC",
    seurat_clusters == 3  ~ "EC",
    seurat_clusters == 4  ~ "PT",
    seurat_clusters == 5  ~ "PT",
    seurat_clusters == 6  ~ "PODO",
    seurat_clusters == 7  ~ "EC",
    seurat_clusters == 8  ~ "IMM",
    seurat_clusters == 9  ~ "PT",
    seurat_clusters == 10 ~ "MES",
    seurat_clusters == 11 ~ "Doublet_1",
    seurat_clusters == 12 ~ "IMM",
    seurat_clusters == 13 ~ "DCT",
    seurat_clusters == 14 ~ "IMM",
    seurat_clusters == 15 ~ "PC",
    seurat_clusters == 16 ~ "IMM",
    seurat_clusters == 17 ~ "IC",
    seurat_clusters == 18 ~ "IMM",
    seurat_clusters == 19 ~ "PEC",
    seurat_clusters == 20 ~ "EC",
    seurat_clusters == 21 ~ "IMM",
    seurat_clusters == 22 ~ "TAL",
    seurat_clusters == 23 ~ "Doublet_2",
    ))

#SO2@meta.data$class.Merge <- factor(SO2@meta.data$class.Merge , levels = c("Fibroblasts", "Medullary Fibroblasts", "Proliferating", "Mesangial Cells", "Pericytes", "Efferent #Artery", "Afferent Artery", "JGA"))

DimPlot(object = SO2, reduction = "umap", group.by = "class_merge", label = TRUE)

```

## Keep Glom Cell Types

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

Idents(SO2) <- SO2@meta.data$class_merge

keep <- c("PODO", "EC", "MES", "IMM", "PEC", "VSMC")

SO3 <- subset(SO2, subset = class_merge %in% keep)

DimPlot(object = SO3, reduction = "umap", group.by = "class_merge", label = TRUE)
```

# Recluster #2

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

SO4 <- SCTransform(SO3) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = 1) %>%
    RunUMAP(dims = 1:30)

```

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

DimPlot(SO4, reduction = "umap", label = TRUE)

DimPlot(SO4, reduction = "umap", group.by = "sample")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=12}

DimPlot(SO4, reduction = "umap", split.by = "sample")

```

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}


DimPlot(SO4, reduction = "umap", group.by = "subclass_merge", label = TRUE)

DimPlot(SO4, reduction = "umap", group.by = "class_merge", label = TRUE)

```

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center',fig.width=12, fig.height=10}

DotPlot(SO4,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```

# Clean DotPlot

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

SO4 <- subset(SO4, subset = seurat_clusters == c("17"), invert = TRUE)

Idents(SO4) <- SO4@meta.data$class_merge

SO4@meta.data$class_merge <- factor(SO4@meta.data$class_merge , levels = c("PODO", "EC", "MES", "IMM", "PEC", "VSMC"))

Idents(SO4) <- SO4@meta.data$class_merge

markers.to.plot2 <- c("Nphs2",        # Podo
                      "Flt1",         # Endo
                      "Pdgfrb",       # Perivascular
                      "Ptprc",        # Immune
                      "Ncam1",        # PEC
                      "Acta2"        # Mural
)                      
                      
DotPlot(SO4,
features = markers.to.plot2,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

for (i in markers.to.plot2)
{
  print(FeaturePlot(SO4, features = i, label = TRUE))
}



```

# Save Object

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(SO4, here("outputs", "Alports_4M.rds"))

```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

