---
title: "Alports Glom Mes Analysis"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("org.Mm.eg.db")


library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)
library(kableExtra) 


options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```

# Introduction

## Email from Sargis on 11/3/2025

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- readRDS(file = here("outputs", "Alports_041525.rds"))

DimPlot(SO, label = T)

```

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}

SO_mes <- subset(SO, idents = "MES")

DimPlot(SO_mes, label = T)

SO_mes <- SCTransform(SO_mes) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = 1) %>%
    RunUMAP(dims = 1:30)

DimPlot(SO_mes, label = T)

DimPlot(SO_mes, label = T, group.by = "group")

FeaturePlot(SO_mes, "Nphs2")

VlnPlot(SO_mes, "Nphs2")

FeaturePlot(SO_mes, "Kdr")

VlnPlot(SO_mes, "Kdr")

FeaturePlot(SO_mes, "Pdgfrb")

VlnPlot(SO_mes, "Pdgfrb")

```


```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}

SO_mes <- subset(SO_mes, idents = c("3", "4", "7"), invert = TRUE)

DimPlot(SO_mes)

```


## Heatmap

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.height=10}

SO.markers <- FindAllMarkers(SO_mes, only.pos = TRUE, min.pct = .25)

SO.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top10

DoHeatmap(SO_mes, features = top10$gene) + NoLegend() +
  theme(plot.margin = margin(20, 80, 10, 10))

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

t1 <- table(SO_mes@meta.data$seurat_clusters, SO_mes$group)[, c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp")]
#t1

prop.t1 <- prop.table(t1, margin = 2) 
#prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

library(scales)   # ggplot2 depends on this

# ggplot2 default discrete palette (6 colors)
my_colors2 <- scales::hue_pal()(12)


ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  scale_fill_manual(values = my_colors2) +
  theme_classic()

```

#### DEG Setup

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Condition_Baseline <- "WT"

Condition_A <- "4M_Alp"
  
Condition_B <- "4M_Alp_1EV"

Condition_C <- "4M_Alp_2EV"

Condition_D <- "6M_Alp"

clusters <- unique(SO_mes@meta.data$class_merge)
names(clusters) <- clusters

SO <- PrepSCTFindMarkers(SO_mes)
DEG.env <- new.env()

Idents(SO_mes) <- SO_mes@meta.data$group

DefaultAssay(SO_mes) <- "SCT"

```

#### df1

4M_Alp vs. WT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DEGs <- FindMarkers(SO_mes, ident.1 = Condition_A, ident.2 = Condition_Baseline, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df1 <- DEGs


```

#### df2

4M_Alp_1EV vs. 4M_Alp

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DEGs <- FindMarkers(SO_mes, ident.1 = Condition_B, ident.2 = Condition_A, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df2 <- DEGs


```

#### df3

4M_Alp_2EV vs. 4M_Alp

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DEGs <- FindMarkers(SO_mes, ident.1 = Condition_C, ident.2 = Condition_A, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df3 <- DEGs

```

#### df4

Corrected from original version on 3.6.26 (orginal comparison was 6M_Alp vs. WT...this is moved to df5)

6M_Alp vs. 4M_Alp

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DEGs <- FindMarkers(SO_mes, ident.1 = Condition_D, ident.2 = Condition_A, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df4 <- DEGs

```

#### df5

6M_Alp vs. WT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DEGs <- FindMarkers(SO_mes, ident.1 = Condition_D, ident.2 = Condition_Baseline, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df5 <- DEGs

```


#### Write Spreadsheet of DEG Lits

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# Create workbook
wb <- createWorkbook()

# Add sheets + write data
addWorksheet(wb, "df1 - 4M_Alp vs. WT")
writeData(wb, "df1 - 4M_Alp vs. WT", df1)

addWorksheet(wb, "df2 - 4M_Alp_1EV vs. 4M_Alp")
writeData(wb, "df2 - 4M_Alp_1EV vs. 4M_Alp", df2)

addWorksheet(wb, "df3 - 4M_Alp_2EV vs. 4M_Alp")
writeData(wb, "df3 - 4M_Alp_2EV vs. 4M_Alp", df3)

addWorksheet(wb, "df4 - 6M_Alp vs. 4M_Alp")
writeData(wb, "df4 - 6M_Alp vs. 4M_Alp", df4)

addWorksheet(wb, "df5 - 6M_Alp vs. WT")
writeData(wb, "df5 - 6M_Alp vs. WT", df5)

# Save workbook
saveWorkbook(wb, "MES_compiled_DEG_lists_042926.xlsx", overwrite = TRUE)

```

#### Filter and Split Positive and Negative Lists

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df_summary <- tibble(
  dataframe = c("df1", "df2", "df3", "df4", "df5"),
  nrows = c(nrow(df1), nrow(df2), nrow(df3), nrow(df4), nrow(df5))
)

# Plot
ggplot(df_summary, aes(x = dataframe, y = nrows, fill = dataframe)) +
  geom_col() +
  labs(
    title = "Number of DEGs",
    x = "Dataframe",
    y = "Number of DEGs"
  ) +
  theme_classic() +
  theme(legend.position = "none")




```


#### Filter and Split Positive and Negative Lists

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1_pos <- df1 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df1_pos

df1_neg <- df1 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df1_neg

df2_pos <- df2 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df2_pos

df2_neg <- df2 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df2_neg

df3_pos <- df3 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df3_pos

df3_neg <- df3 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df3_neg

df4_pos <- df4 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df4_pos

df4_neg <- df4 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df4_neg

df5_pos <- df5 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df5_pos

df5_neg <- df5 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df5_neg

# List only the pos/neg data frames
dfs <- list(
  df1_pos = df1_pos,
  df1_neg = df1_neg,
  df2_pos = df2_pos,
  df2_neg = df2_neg,
  df3_pos = df3_pos,
  df3_neg = df3_neg,
  df4_pos = df4_pos,
  df4_neg = df4_neg,
  df5_pos = df5_pos,
  df5_neg = df5_neg
)

# Create summary table
df_summary <- tibble(
  dataframe = names(dfs),
  nrows = sapply(dfs, nrow),
  group = sub("_(pos|neg)$", "", names(dfs)),  # extract df1, df2, df3, df4
  type = sub("^df[0-9]+_", "", names(dfs))     # extract pos / neg
)

# Plot grouped bars
ggplot(df_summary, aes(x = group, y = nrows, fill = type)) +
  geom_col(position = "dodge") +
  labs(
    title = "Number of DEGs",
    x = "Dataframe",
    y = "DEGs",
    fill = "Subset"
  ) +
  theme_classic()

```

### Improved Graphs on 030626

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


df_summary <- tibble(
  dataframe = c("mid-AS", "late-AS"),
  nrows = c(nrow(df1), nrow(df5))
)

df_summary <- df_summary %>%
  mutate(dataframe = factor(dataframe, levels = c("mid-AS", "late-AS")))

# Plot
ggplot(df_summary, aes(x = dataframe, y = nrows, fill = dataframe)) +
  geom_col() +
  labs(
    title = NULL,
    x = NULL,
    y = "Number of DEGs"
  ) +
  theme_classic() +
  theme(
    axis.ticks = element_line(size = 2),   # 1.2 * 3
    axis.line = element_line(size = 2),    # 1.2 * 3
    axis.title = element_text(size = 24),    # 20 * 2
    axis.text = element_text(size = 24),     # 16 * 2
    legend.position = "none"
  )

```


### Improved Graphs on 030626

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

dfs <- list(
  df1_pos = df1_pos,
  df1_neg = df1_neg,
  df5_pos = df5_pos,
  df5_neg = df5_neg
)

# Create summary table
df_summary2 <- tibble(
  dataframe = names(dfs),
  nrows = sapply(dfs, nrow),
  group = sub("_(pos|neg)$", "", names(dfs)),  # extract df1, df2, df3, df4
  type = sub("^df[0-9]+_", "", names(dfs))     # extract pos / neg
)

df_summary2 <- df_summary2 %>%
  mutate(group = case_when(
    group == "df1" ~ "mid-AS",
    group == "df5" ~ "late-AS",
    TRUE ~ group
  ))

df_summary2 <- df_summary2 %>%
  mutate(group = factor(group, levels = c("mid-AS", "late-AS")))

# Plot grouped bars
ggplot(df_summary2, aes(x = group, y = nrows, fill = type)) +
  geom_col(position = "dodge") +
  labs(
    title = NULL,
    x = NULL,
    y = "DEGs",
    fill = "Subset"
  ) +
  theme_classic() +
  theme(
    axis.ticks = element_line(size = 2),
    axis.line = element_line(size = 2),
    axis.title = element_text(size = 24),
    axis.text = element_text(size = 24),
    legend.title = element_text(size = 24),
    legend.text = element_text(size = 24)
  )

```
# Q4

## Pseudobulk PCA

### ALL GROUPS

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# head(SO_mes@meta.data)

pbo <- AggregateExpression(SO_mes, 
                           group.by = "sample",
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

pbo <- pbo$RNA
# pbo[1:5, 1:5]
colnames(pbo) <- sub("^g", "", colnames(pbo))
# colnames(pbo)[1:5]

# DimPlot(SO_mes, group.by = "sample")

metadata <- SO_mes@meta.data %>%
  as.data.frame() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(
    Alports = names(sort(table(alports), decreasing = TRUE))[1],  # Most common value
    Batch = names(sort(table(batch), decreasing = TRUE))[1],
    Group = names(sort(table(group), decreasing = TRUE))[1]
  ) %>%
  as.data.frame()

rownames(metadata) <- metadata$sample

# head(metadata)

dds <- DESeqDataSetFromMatrix(countData = round(pbo),
                              colData = metadata,
                              design = ~ sample)

rld <- rlog(dds, blind=TRUE)

sample_names <- unique(rld$sample)

```

#### PCA By **Sample**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


# Plot PCA
sample<-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "sample") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(sample)

```

#### PCA By **Alports**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


alports <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Alports") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(alports)

```

#### PCA By **Group**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


group <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Group") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(group)


```

### No 6M Alports

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# head(SO_mes@meta.data)

pbo <- AggregateExpression(SO_mes, 
                           group.by = "sample",
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

pbo <- pbo$RNA
# pbo[1:5, 1:5]
colnames(pbo) <- sub("^g", "", colnames(pbo))
pbo <- pbo[, !colnames(pbo) %in% c("A", "B", "C")]
#colnames(pbo)

# DimPlot(SO_mes, group.by = "sample")

metadata <- SO_mes@meta.data %>%
  as.data.frame() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(
    Alports = names(sort(table(alports), decreasing = TRUE))[1],  # Most common value
    Batch = names(sort(table(batch), decreasing = TRUE))[1],
    Group = names(sort(table(group), decreasing = TRUE))[1]
  ) %>%
  as.data.frame()

metadata <- metadata %>% 
  dplyr::filter(Group != "6M_Alp")

rownames(metadata) <- metadata$sample

# head(metadata)

dds <- DESeqDataSetFromMatrix(countData = round(pbo),
                              colData = metadata,
                              design = ~ sample)

rld <- rlog(dds, blind=TRUE)

sample_names <- unique(rld$sample)

```

#### PCA By **Sample**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


# Plot PCA
sample<-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "sample") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(sample)

```

#### PCA By **Alports**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


alports <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Alports") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(alports)

```

#### PCA By **Group**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


group <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Group") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(group)


```


### No EVs

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# head(SO_mes@meta.data)

pbo <- AggregateExpression(SO_mes, 
                           group.by = "sample",
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

pbo <- pbo$RNA
# pbo[1:5, 1:5]
colnames(pbo) <- sub("^g", "", colnames(pbo))
pbo <- pbo[, !colnames(pbo) %in% c("473", "375", "364", "368", "475", "362")]
#colnames(pbo)

# DimPlot(SO_mes, group.by = "sample")

metadata <- SO_mes@meta.data %>%
  as.data.frame() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(
    Alports = names(sort(table(alports), decreasing = TRUE))[1],  # Most common value
    Batch = names(sort(table(batch), decreasing = TRUE))[1],
    Group = names(sort(table(group), decreasing = TRUE))[1]
  ) %>%
  as.data.frame()

metadata <- metadata %>% 
  dplyr::filter(Group != "4M_Alp_1EV")

metadata <- metadata %>% 
  dplyr::filter(Group != "4M_Alp_2EV")

rownames(metadata) <- metadata$sample

# head(metadata)

dds <- DESeqDataSetFromMatrix(countData = round(pbo),
                              colData = metadata,
                              design = ~ sample)

rld <- rlog(dds, blind=TRUE)

sample_names <- unique(rld$sample)

```

#### PCA By **Sample**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


# Plot PCA
sample<-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "sample") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(sample)

```

#### PCA By **Alports**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


alports <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Alports") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(alports)

```

#### PCA By **Group**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


group <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Group") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(group)


```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

