---
title: "Alports Glom EC Analysis"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("org.Mm.eg.db")


library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)
library(kableExtra) 


options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```

# Introduction

## Email from Sargis on 11/3/2025

Q1: The color codes assigned to GEC subtypes in the heatmap do not match those used in the bar graphs showing subtype frequences. I was able to classify the subtypes based on the top 5 markers in the heatmap but was not able to compare between the groups due to this issue. 

Q2: EC subtype classification: Please, let me know what you think about it. 

Q3: Could you please help me identify the genes in these Venn diagrams, (also including the comparison between 1EV/AS vs 4M/WT) especially those overlapping. I tried to extract them from the excel files using different cut offs, but I kept getting different numbers than yours. Could not figure out why. And then it would be interesting to do pathway analysis (or GO analysis) on these sets of genes for each Venn diagram. 

Q4: Would it be possible to have a pseudobulk PCA analysis of the groups? like for examples WT, 4M, 1EV and 2EV in one PCA plot, and separately WT, 4M AS and 6m AS in a different PCA plot?

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- readRDS(file = here("outputs", "Alports_Glom_EC.rds"))

# DimPlot(SO, label = T)

```
# Q1

### Creating a Consistent Color Palette v1

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

library(scales)   # ggplot2 depends on this

# ggplot2 default discrete palette (6 colors)
my_colors <- scales::hue_pal()(7)

# my_colors

type_levels <- c(
  "Type_A",
  "Type_B",
  "Type_C",
  "Type_D",
  "Type_E",
  "Type_Mito",
  "Type_Prolif"
)

names(my_colors) <- type_levels
#my_colors

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- PrepSCTFindMarkers(SO)

SO.markers <- FindAllMarkers(SO, only.pos = TRUE, min.pct = .25)

SO.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top10

# DoHeatmap(SO, features = top10$gene) + NoLegend()

```

## Rename v1

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results='hide'}


SO2 <- SO

SO2@meta.data <- SO2@meta.data %>% 
  mutate(subclass_EC = dplyr::case_when(
    seurat_clusters == 0  ~ "Type_A",
    seurat_clusters == 1  ~ "Type_A",
    seurat_clusters == 2  ~ "Type_B",
    seurat_clusters == 3  ~ "Type_B",
    seurat_clusters == 4  ~ "Type_C",
    seurat_clusters == 5  ~ "Type_D",
    seurat_clusters == 6  ~ "Type_Mito",
    seurat_clusters == 7  ~ "Type_D",
    seurat_clusters == 8  ~ "Type_E",
    seurat_clusters == 9  ~ "Type_Prolif"
  ))

SO2@meta.data$subclass_EC <- factor(
  SO2@meta.data$subclass_EC,
  levels = c(
    "Type_A",
    "Type_B",
    "Type_C",
    "Type_D",
    "Type_E",
    "Type_Mito",
    "Type_Prolif"
  )
)
```

### Creating a Consistent Color Palette v1

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

library(scales)   # ggplot2 depends on this

# ggplot2 default discrete palette (6 colors)
my_colors <- scales::hue_pal()(7)

# my_colors

type_levels <- c(
  "Type_A",
  "Type_B",
  "Type_C",
  "Type_D",
  "Type_E",
  "Type_Mito",
  "Type_Prolif"
)

names(my_colors) <- type_levels
#my_colors

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DimPlot(SO2, group.by = "subclass_EC",) + scale_color_manual(values = my_colors)

```

## Rename v2

### Creating a Consistent Color Palette v2

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

library(scales)   # ggplot2 depends on this

# ggplot2 default discrete palette (6 colors)
my_colors2 <- scales::hue_pal()(7)

# my_colors

type_levels <- c(
  "Type_1",
  "Type_2",
  "Type_3",
  "Type_4",
  "Type_5",
  "Type_6",
  "Type_7"
)

names(my_colors2) <- type_levels
my_colors2

```

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results='hide'}

SO2@meta.data <- SO2@meta.data %>% 
  mutate(subclass_EC_2 = dplyr::case_when(
    subclass_EC == "Type_A"  ~ "Type_1",
    subclass_EC == "Type_B"  ~ "Type_2",
    subclass_EC == "Type_C"  ~ "Type_4",
    subclass_EC == "Type_D"  ~ "Type_3",
    subclass_EC == "Type_E" ~ "Type_6",
    subclass_EC == "Type_Mito"  ~ "Type_5",
    subclass_EC == "Type_Prolif"  ~ "Type_7"
  ))

SO2@meta.data$subclass_EC_2 <- factor(
  SO2@meta.data$subclass_EC_2,
  levels = c(
    "Type_1",
    "Type_2",
    "Type_3",
    "Type_4",
    "Type_5",
    "Type_6",
    "Type_7"
  )
)

DimPlot(SO2, group.by = "subclass_EC_2") + scale_color_manual(values = my_colors2)

```

## Table of EC Cell Type Identities

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=10}

cell_data <- data.frame(
  Cluster = c("Type_1", "Type_3", "Type_4", "Type_5", "Type_2", "Type_6", "Type_7"),
  Markers = c(
    "Csrnp1, Nfkbiz, Ppp1r10, Socs1, Dusp5",
    "Slc6a6, Pltp, Atp13a3, Aqp1, Slc26a10",
    "Slc45a4, Gja5, Esm1, Cxcr4, Hey1",
    "mt-Nd5, mt-Nd4, mt-Nd3, mt-Nd2, mt-Nd4l",
    "Ddnrb, Fabp4, Gda, Akap12, Pde7b",
    "Dlk1, Rtl1, Sez6l2, Egln3",
    "Uhrf1, Gins2, Mki67, Fignl1, Rad51"
  ),
  Cell_State = c(
    "Immune-regulatory",
    "Transport-focused",
    "Angiogenic",
    "Mitochondrial-rich",
    "Lipid, structural maintenance, quiescent",
    "Developmental",
    "Proliferative"
  ),
  stringsAsFactors = FALSE
)

# Reorder numerically by the number after "Type_"
cell_data <- cell_data %>%
  mutate(type_num = as.numeric(str_extract(Cluster, "\\d+"))) %>%
  arrange(type_num) %>%
  select(-type_num)  # remove helper column

# Render table with kable
kable(cell_data, "html", align = c("c","c","c"), caption = "<b><span style='font-size:16pt'>Cell Clusters with Marker Genes and Cell States</span></b>") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  row_spec(0, bold = TRUE, color = "black") # bold header

```

## Heatmap

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=12}

Idents(SO2) <- SO2@meta.data$subclass_EC_2

SO.markers <- FindAllMarkers(SO2, only.pos = TRUE, min.pct = .25)

SO.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top10

DoHeatmap(SO2, features = top10$gene) + NoLegend() +
  theme(plot.margin = margin(20, 80, 10, 10))

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

t1 <- table(SO2@meta.data$subclass_EC_2, SO2$group)[, c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp")]
#t1

prop.t1 <- prop.table(t1, margin = 2) 
#prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  scale_fill_manual(values = my_colors2) +
  theme_classic()

```

# Q2

Seems good to me and see above for changing the names of the EC clusters and creating a consistnent color palette. 

# Q3

## The Problem

Part of the problem you were facing in trying to **compare the Venn Diagrams and the CSV DEG lists** were created in different ways and from different files at different points in the analysis. 

DEG lists:

From **Alports_Quantify_Number_of_DEGs_WTv4M_Alp_1EV_09182025**

Using **SCT** assay and **Seurat_5.3.0**

for (i in clusters) {
DefaultAssay(SO) <- "SCT"
DEGs <- FindMarkers(SO, ident.1 = "4M_Alp_1EV", ident.2 = "WT", group.by = "group" , subset.ident = i, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]
assign(paste0("Alports_4M_",names(clusters[i])), DEGs, envir = DEG.env)

Venn Diagram: **Alports_subset_ECs_Quantify_Reversal_Genes_0528202**

I think using **RNA** assay and **Seurat_4.4.0**

df1 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_A, ident.2 = Condition_Baseline, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

## The Solution

Start with the DEG lists from **Alports_Quantify_Number_of_DEGs_WTv4M_Alp_1EV_09182025** and create Venn Diagrams and new intersection lists to perform Pathways Analysis.

## Update: **Solution Didn't work**

### The Previous DEG lists aren't the correct comparisons for making the Venn Diagrams
### They were really setup to look at all the different populations...not just ECs

### Solution #2

Recreate the DEG lists again from the **Alports_Glom_EC.rds** dataset and save them as csv files as well. 

### Re-create DEG Lists

df1 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_A, ident.2 = Condition_Baseline, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df2 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_B, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df3 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_C, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df4 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_D, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

#### DEG Setup

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Condition_Baseline <- "WT"

Condition_A <- "4M_Alp"
  
Condition_B <- "4M_Alp_1EV"

Condition_C <- "4M_Alp_2EV"

Condition_D <- "6M_Alp"

clusters <- unique(SO2@meta.data$class_merge)
names(clusters) <- clusters

SO <- PrepSCTFindMarkers(SO2)
DEG.env <- new.env()

Idents(SO2) <- SO2@meta.data$group

DefaultAssay(SO2) <- "SCT"

```

#### df1

4M_Alp vs. WT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DEGs <- FindMarkers(SO2, ident.1 = Condition_A, ident.2 = Condition_Baseline, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df1 <- DEGs


```

#### df2

4M_Alp_1EV vs. 4M_Alp

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


DEGs <- FindMarkers(SO2, ident.1 = Condition_B, ident.2 = Condition_A, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df2 <- DEGs


```

#### df3

4M_Alp_2EV vs. 4M_Alp

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DEGs <- FindMarkers(SO2, ident.1 = Condition_C, ident.2 = Condition_A, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df3 <- DEGs

```

#### df4

6M_Alp vs. WT

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

DEGs <- FindMarkers(SO2, ident.1 = Condition_D, ident.2 = Condition_Baseline, logfc.threshold = .1, min.pct = .1)
DEGs <- DEGs %>% rownames_to_column("gene") %>% filter(p_val_adj < .05)
DEGs <- DEGs[order(DEGs[,"avg_log2FC"], decreasing = TRUE),]

#remove the first column of from the dataframe DEGs
DEGs <- rownames_to_column(DEGs)
DEGs <- DEGs[,-1]

df <- head(DEGs, 20)

df

df4 <- DEGs

```

#### Write Spreadsheet of DEG Lits

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# Create workbook
wb <- createWorkbook()

# Add sheets + write data
addWorksheet(wb, "df1 - 4M_Alp vs. WT")
writeData(wb, "df1 - 4M_Alp vs. WT", df1)

addWorksheet(wb, "df2 - 4M_Alp_1EV vs. 4M_Alp")
writeData(wb, "df2 - 4M_Alp_1EV vs. 4M_Alp", df2)

addWorksheet(wb, "df3 - 4M_Alp_2EV vs. 4M_Alp")
writeData(wb, "df3 - 4M_Alp_2EV vs. 4M_Alp", df3)

addWorksheet(wb, "df4 - 6M_Alp vs. 4M_Alp")
writeData(wb, "df4 - 6M_Alp vs. 4M_Alp", df4)

# Save workbook
saveWorkbook(wb, "compiled_DEG_lists.xlsx", overwrite = TRUE)

```

#### Filter and Split Positive and Negative Lists

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df_summary <- tibble(
  dataframe = c("df1", "df2", "df3", "df4"),
  nrows = c(nrow(df1), nrow(df2), nrow(df3), nrow(df4))
)

# Plot
ggplot(df_summary, aes(x = dataframe, y = nrows, fill = dataframe)) +
  geom_col() +
  labs(
    title = "Number of DEGs",
    x = "Dataframe",
    y = "Number of DEGs"
  ) +
  theme_classic() +
  theme(legend.position = "none")

```


#### Filter and Split Positive and Negative Lists

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1_pos <- df1 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df1_pos

df1_neg <- df1 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df1_neg

df2_pos <- df2 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df2_pos

df2_neg <- df2 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df2_neg

df3_pos <- df3 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df3_pos

df3_neg <- df3 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df3_neg

df4_pos <- df4 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC > .25) %>% arrange(desc(abs(avg_log2FC)))

df4_pos

df4_neg <- df4 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% filter(avg_log2FC < -.25) %>% arrange(desc(abs(avg_log2FC)))

df4_neg

# List only the pos/neg data frames
dfs <- list(
  df1_pos = df1_pos,
  df1_neg = df1_neg,
  df2_pos = df2_pos,
  df2_neg = df2_neg,
  df3_pos = df3_pos,
  df3_neg = df3_neg,
  df4_pos = df4_pos,
  df4_neg = df4_neg
)

# Create summary table
df_summary <- tibble(
  dataframe = names(dfs),
  nrows = sapply(dfs, nrow),
  group = sub("_(pos|neg)$", "", names(dfs)),  # extract df1, df2, df3, df4
  type = sub("^df[0-9]+_", "", names(dfs))     # extract pos / neg
)

# Plot grouped bars
ggplot(df_summary, aes(x = group, y = nrows, fill = type)) +
  geom_col(position = "dodge") +
  labs(
    title = "Number of DEGs",
    x = "Dataframe",
    y = "DEGs",
    fill = "Subset"
  ) +
  theme_classic()

```

## Venn Diagram Readout

### Up -> Down

#### 4M vs 4M_1EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- df1_pos$gene
set2 <- df2_neg$gene
venn <- ggvenn(list('Up in 4M' = set1, 'Down in 1EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = FALSE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_1EV <- test$'A:B'

# Find the maximum length
max_len <- max(length(test$A), length(test$B), length(test$`A:B`))

# Pad each vector to the same length
A_col  <- c(test$A,  rep(NA, max_len - length(test$A)))
B_col  <- c(test$B,  rep(NA, max_len - length(test$B)))
AB_col <- c(test$`A:B`, rep(NA, max_len - length(test$`A:B`)))

# Create a new dataframe
venn_df <- data.frame(
  'Up_in_4M' = A_col,
  'Down_in_1EV' = B_col,
  'Intersection' = AB_col,
  stringsAsFactors = FALSE
)

# Create workbook
wb <- createWorkbook()

# Add a worksheet
addWorksheet(wb, "venn_df")

# Write the data
writeData(wb, "venn_df", venn_df)

# Save the Excel file
saveWorkbook(wb, "venn_gene_lists_4M_UP_and_4M1EV_DOWN.xlsx", overwrite = TRUE)

```

#### Pathways

##### Up_in_4M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = A_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_4M1EV_DOWN_x_Up_in_4M.xlsx")

```

##### Down_in_1EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = B_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_4M1EV_DOWN_x_Down_in_1EV.xlsx")

```

##### INTERSECTION

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = AB_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_4M1EV_DOWN_x_INTERSECTION.xlsx")

```

#### 4M vs 4M_2EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- df1_pos$gene
set2 <- df3_neg$gene
venn <- ggvenn(list('Up in 4M' = set1, 'Down in 2EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = FALSE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_2EV <- test$'A:B'

# Find the maximum length
max_len <- max(length(test$A), length(test$B), length(test$`A:B`))

# Pad each vector to the same length
A_col  <- c(test$A,  rep(NA, max_len - length(test$A)))
B_col  <- c(test$B,  rep(NA, max_len - length(test$B)))
AB_col <- c(test$`A:B`, rep(NA, max_len - length(test$`A:B`)))

# Create a new dataframe
venn_df <- data.frame(
  'Up_in_4M' = A_col,
  'Down_in_2EV' = B_col,
  'Intersection' = AB_col,
  stringsAsFactors = FALSE
)

# Create workbook
wb <- createWorkbook()

# Add a worksheet
addWorksheet(wb, "venn_df")

# Write the data
writeData(wb, "venn_df", venn_df)

# Save the Excel file
saveWorkbook(wb, "venn_gene_lists_4M_UP_and_4M2EV_DOWN.xlsx", overwrite = TRUE)

```

#### Pathways

##### Up_in_4M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = A_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_4M2EV_DOWN_x_Up_in_4M.xlsx")

```

##### Down_in_2EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = B_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_4M2EV_DOWN_x_Down_in_2EV.xlsx")

```

##### INTERSECTION

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = AB_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_4M2EV_DOWN_x_INTERSECTION.xlsx")

```


#### 4M vs 6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- df1_pos$gene
set2 <- df4_neg$gene
venn <- ggvenn(list('Up in 4M' = set1, 'Down in 6M' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = FALSE)

test <- attributes(ItemsList)$intersections

genelist_6M_Alp <- test$'A:B'

# Find the maximum length
max_len <- max(length(test$A), length(test$B), length(test$`A:B`))

# Pad each vector to the same length
A_col  <- c(test$A,  rep(NA, max_len - length(test$A)))
B_col  <- c(test$B,  rep(NA, max_len - length(test$B)))
AB_col <- c(test$`A:B`, rep(NA, max_len - length(test$`A:B`)))

# Create a new dataframe
venn_df <- data.frame(
  'Up_in_4M' = A_col,
  'Down in 6M' = B_col,
  'Intersection' = AB_col,
  stringsAsFactors = FALSE
)

# Create workbook
wb <- createWorkbook()

# Add a worksheet
addWorksheet(wb, "venn_df")

# Write the data
writeData(wb, "venn_df", venn_df)

# Save the Excel file
saveWorkbook(wb, "venn_gene_lists_4M_UP_and_6M_DOWN.xlsx", overwrite = TRUE)

```

#### Pathways

##### Up_in_4M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = A_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_6M_DOWN_x_Up_in_4M.xlsx")

```

##### Down_in_6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = B_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_6M_DOWN_x_Down_in_6M.xlsx")

```

##### INTERSECTION

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = AB_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_UP_and_6M_DOWN_x_INTERSECTION.xlsx")

```


### Down -> Up

#### 4M vs 4M_1EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- df1_neg$gene
set2 <- df2_pos$gene
venn <- ggvenn(list('Down in 4M' = set1, 'Up in 1EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = FALSE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_1EV <- test$'A:B'

# Find the maximum length
max_len <- max(length(test$A), length(test$B), length(test$`A:B`))

# Pad each vector to the same length
A_col  <- c(test$A,  rep(NA, max_len - length(test$A)))
B_col  <- c(test$B,  rep(NA, max_len - length(test$B)))
AB_col <- c(test$`A:B`, rep(NA, max_len - length(test$`A:B`)))

# Create a new dataframe
venn_df <- data.frame(
  'Down in 4M' = A_col,
  'Up in 1EV' = B_col,
  'Intersection' = AB_col,
  stringsAsFactors = FALSE
)

# Create workbook
wb <- createWorkbook()

# Add a worksheet
addWorksheet(wb, "venn_df")

# Write the data
writeData(wb, "venn_df", venn_df)

# Save the Excel file
saveWorkbook(wb, "venn_gene_lists_4M_DOWN_and_4M1EV_UP.xlsx", overwrite = TRUE)

```

#### Pathways

##### Down_in_4M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = A_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_4M1EV_UP_x_Down_in_4M.xlsx")

```

##### Up_in_6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = B_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_4M1EV_UP_x_Up_in_1EV.xlsx")

```

##### INTERSECTION

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = AB_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_4M1EV_UP_x_INTERSECTION.xlsx")

```

#### 4M vs 4M_2EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- df1_neg$gene
set2 <- df3_pos$gene
venn <- ggvenn(list('Down in 4M' = set1, 'Up in 2EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = FALSE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_2EV <- test$'A:B'

# Find the maximum length
max_len <- max(length(test$A), length(test$B), length(test$`A:B`))

# Pad each vector to the same length
A_col  <- c(test$A,  rep(NA, max_len - length(test$A)))
B_col  <- c(test$B,  rep(NA, max_len - length(test$B)))
AB_col <- c(test$`A:B`, rep(NA, max_len - length(test$`A:B`)))

# Create a new dataframe
venn_df <- data.frame(
  'Down in 4M' = A_col,
  'Up in 2EV' = B_col,
  'Intersection' = AB_col,
  stringsAsFactors = FALSE
)

# Create workbook
wb <- createWorkbook()

# Add a worksheet
addWorksheet(wb, "venn_df")

# Write the data
writeData(wb, "venn_df", venn_df)

# Save the Excel file
saveWorkbook(wb, "venn_gene_lists_4M_DOWN_and_4M2EV_UP.xlsx", overwrite = TRUE)

```

#### Pathways

##### Down_in_4M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = A_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_4M2EV_UP_x_Down_in_4M.xlsx")

```

##### Up_in_4M_2EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = B_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_4M2EV_UP_x_Up_in_2EV.xlsx")

```

##### INTERSECTION

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = AB_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_4M2EV_UP_x_INTERSECTION.xlsx")

```

#### 4M vs 6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- df1_neg$gene
set2 <- df4_pos$gene
venn <- ggvenn(list('Down in 4M' = set1, 'Up in 6M' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = FALSE)

test <- attributes(ItemsList)$intersections

genelist_6M_Alp <- test$'A:B'

# Find the maximum length
max_len <- max(length(test$A), length(test$B), length(test$`A:B`))

# Pad each vector to the same length
A_col  <- c(test$A,  rep(NA, max_len - length(test$A)))
B_col  <- c(test$B,  rep(NA, max_len - length(test$B)))
AB_col <- c(test$`A:B`, rep(NA, max_len - length(test$`A:B`)))

# Create a new dataframe
venn_df <- data.frame(
  'Down in 4M' = A_col,
  'Up in 6M' = B_col,
  'Intersection' = AB_col,
  stringsAsFactors = FALSE
)

# Create workbook
wb <- createWorkbook()

# Add a worksheet
addWorksheet(wb, "venn_df")

# Write the data
writeData(wb, "venn_df", venn_df)

# Save the Excel file
saveWorkbook(wb, "venn_gene_lists_4M_DOWN_and_6M_UP.xlsx", overwrite = TRUE)

```

#### Pathways

##### Down_in_4M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = A_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_6M_UP_x_Down_in_4M.xlsx")

```

##### Up_in_6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = B_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_6M_UP_x_Up_in_6M.xlsx")

```

##### INTERSECTION

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

ENTREZ_list <- bitr(geneID = AB_col,   #input gene id
                    fromType = "SYMBOL",           #input id type
                    toType = "ENTREZID",           #output id type
                    OrgDb = "org.Mm.eg.db"         #annotation Db
                    )

go <- enrichGO(gene = ENTREZ_list$ENTREZID,           #a vector of entrez gene id
                   OrgDb = "org.Mm.eg.db",    
                   ont = "BP",
                   readable = TRUE)              #whether mapping gene ID to gene Name
go

dotplot(go) +
    ggtitle("") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme_classic() + 
    scale_x_reverse()  # Reverses the x-axis

df_pathways <- go@result

write.xlsx(df_pathways, file = "pathways_venn_gene_lists_4M_DOWN_and_6M_UP_x_INTERSECTION.xlsx")

```

# Q4

## Pseudobulk PCA

### ALL GROUPS

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# head(SO2@meta.data)

pbo <- AggregateExpression(SO2, 
                           group.by = "sample",
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

pbo <- pbo$RNA
# pbo[1:5, 1:5]
colnames(pbo) <- sub("^g", "", colnames(pbo))
# colnames(pbo)[1:5]

# DimPlot(SO2, group.by = "sample")

metadata <- SO2@meta.data %>%
  as.data.frame() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(
    Alports = names(sort(table(alports), decreasing = TRUE))[1],  # Most common value
    Batch = names(sort(table(batch), decreasing = TRUE))[1],
    Group = names(sort(table(group), decreasing = TRUE))[1]
  ) %>%
  as.data.frame()

rownames(metadata) <- metadata$sample

# head(metadata)

dds <- DESeqDataSetFromMatrix(countData = round(pbo),
                              colData = metadata,
                              design = ~ sample)

rld <- rlog(dds, blind=TRUE)

sample_names <- unique(rld$sample)

```

#### PCA By **Sample**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


# Plot PCA
sample<-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "sample") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(sample)

```

#### PCA By **Alports**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


alports <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Alports") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(alports)

```

#### PCA By **Group**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


group <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Group") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(group)


```

### No 6M Alports

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# head(SO2@meta.data)

pbo <- AggregateExpression(SO2, 
                           group.by = "sample",
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

pbo <- pbo$RNA
# pbo[1:5, 1:5]
colnames(pbo) <- sub("^g", "", colnames(pbo))
pbo <- pbo[, !colnames(pbo) %in% c("A", "B", "C")]
#colnames(pbo)

# DimPlot(SO2, group.by = "sample")

metadata <- SO2@meta.data %>%
  as.data.frame() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(
    Alports = names(sort(table(alports), decreasing = TRUE))[1],  # Most common value
    Batch = names(sort(table(batch), decreasing = TRUE))[1],
    Group = names(sort(table(group), decreasing = TRUE))[1]
  ) %>%
  as.data.frame()

metadata <- metadata %>% 
  dplyr::filter(Group != "6M_Alp")

rownames(metadata) <- metadata$sample

# head(metadata)

dds <- DESeqDataSetFromMatrix(countData = round(pbo),
                              colData = metadata,
                              design = ~ sample)

rld <- rlog(dds, blind=TRUE)

sample_names <- unique(rld$sample)

```

#### PCA By **Sample**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


# Plot PCA
sample<-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "sample") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(sample)

```

#### PCA By **Alports**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


alports <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Alports") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(alports)

```

#### PCA By **Group**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


group <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Group") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(group)


```


### No EVs

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# head(SO2@meta.data)

pbo <- AggregateExpression(SO2, 
                           group.by = "sample",
                           assays = 'RNA', 
                           slot = "counts",
                           return.seurat = FALSE)

pbo <- pbo$RNA
# pbo[1:5, 1:5]
colnames(pbo) <- sub("^g", "", colnames(pbo))
pbo <- pbo[, !colnames(pbo) %in% c("473", "375", "364", "368", "475", "362")]
#colnames(pbo)

# DimPlot(SO2, group.by = "sample")

metadata <- SO2@meta.data %>%
  as.data.frame() %>%
  dplyr::group_by(sample) %>%
  dplyr::summarise(
    Alports = names(sort(table(alports), decreasing = TRUE))[1],  # Most common value
    Batch = names(sort(table(batch), decreasing = TRUE))[1],
    Group = names(sort(table(group), decreasing = TRUE))[1]
  ) %>%
  as.data.frame()

metadata <- metadata %>% 
  dplyr::filter(Group != "4M_Alp_1EV")

metadata <- metadata %>% 
  dplyr::filter(Group != "4M_Alp_2EV")

rownames(metadata) <- metadata$sample

# head(metadata)

dds <- DESeqDataSetFromMatrix(countData = round(pbo),
                              colData = metadata,
                              design = ~ sample)

rld <- rlog(dds, blind=TRUE)

sample_names <- unique(rld$sample)

```

#### PCA By **Sample**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


# Plot PCA
sample<-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "sample") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(sample)

```

#### PCA By **Alports**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


alports <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Alports") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(alports)

```

#### PCA By **Group**

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}


group <-DESeq2::plotPCA(rld, ntop = 1000, intgroup = "Group") + 
  geom_text_repel(aes(label = sample_names)) +
  coord_fixed(ratio = 1.5) +
  theme_classic()

plot(group)


```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

