---
title: "Alports Merging and Filtering"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}

library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)



options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```
# Introduction

## Changes based on meeting with Sargis on 4/22/2025

1) Save Alports_Glom_EC.rds file <br>
2) Change arguments for FindMarkers to `logfc.threshold = 0.1` <br>
3) Write DEG lists to excel file <br>
4) Write Pathways list to excel file <br>
5) Add comparisons of additional groups (WT vs. 6M, 4M vs. 4M_1EV) <br>

## Changes based on meeting with Sargis on 5/28/2025

1) Split glom_EC's by treatment group for split UMAP and Proportion Plot


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- readRDS(file = here("outputs", "Alports_Glom_EC.rds"))

DimPlot(SO, label = T)

```

# FindAllMarkers Heatmap

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO <- PrepSCTFindMarkers(SO)

SO.markers <- FindAllMarkers(SO, only.pos = TRUE, min.pct = .25)

SO.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top10
DoHeatmap(SO, features = top10$gene) + NoLegend()

```

## Rename

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results='hide'}


SO2 <- SO

SO2@meta.data <- SO2@meta.data %>% 
  mutate(subclass_EC = dplyr::case_when(
    seurat_clusters == 0  ~ "Type_A",
    seurat_clusters == 1  ~ "Type_A",
    seurat_clusters == 2  ~ "Type_B",
    seurat_clusters == 3  ~ "Type_B",
    seurat_clusters == 4  ~ "Type_C",
    seurat_clusters == 5  ~ "Type_D",
    seurat_clusters == 6  ~ "Type_Mito",
    seurat_clusters == 7  ~ "Type_D",
    seurat_clusters == 8  ~ "Type_E",
    seurat_clusters == 9  ~ "Type_Prolif"
  ))

DimPlot(SO2, group.by = "subclass_EC")

```


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

t1 <- table(SO2@meta.data$subclass_EC, SO2$group)[, c("WT", "4M_Alp", "4M_Alp_1EV", "4M_Alp_2EV", "6M_Alp")]
t1

prop.t1 <- prop.table(t1, margin = 2) 
prop.t1

t2 <- as.data.frame(t1)
colnames(t2) <- c('Cell_type', 'Group', 'Frequency')

# Original plot

ggplot(t2, aes(fill=Cell_type, y=Frequency, x=Group)) + 
  geom_bar(position="fill", stat = "identity", fun.y = "mean", colour="black") +
  theme_classic()

```

## Identifying EC subtypes

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.width=10}

FeaturePlot(SO2, "Epcam")

FeaturePlot(SO2, "Aqp1")

FeaturePlot(SO2, "mt-Nd2")

VlnPlot(SO2, "percent.mt", group.by = "subclass_EC")

VlnPlot(SO2, "nCount_RNA", group.by = "subclass_EC")

```

## Checking non-EC's

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=10}

Idents(SO2) <- SO2@meta.data$subclass_EC

SO.markers <- FindAllMarkers(SO2, only.pos = TRUE, min.pct = .25)

SO.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top10
DoHeatmap(SO2, features = top10$gene) + NoLegend()

```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

