---
title: "Alports EC: Geneset Quantification"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}

library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)
library(pheatmap)
library(SeuratWrappers)
library(monocle3)

#remotes::install_github('satijalab/seurat-wrappers')

options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()

#usethis::create_github_token()
#gitcreds::gitcreds_set()

#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 

```

## Install and Load Monocle3

https://cole-trapnell-lab.github.io/monocle3/docs/installation/

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}


if (!requireNamespace("BiocManager", quietly = TRUE))
install.packages("BiocManager")
BiocManager::install(version = "3.21")

BiocManager::install(c('BiocGenerics', 'DelayedArray', 'DelayedMatrixStats',
                       'limma', 'lme4', 'S4Vectors', 'SingleCellExperiment',
                       'SummarizedExperiment', 'batchelor', 'HDF5Array',
                       'ggrastr'))

install.packages("devtools")
remotes::install_github("bnprks/BPCells/r")

devtools::install_github('cole-trapnell-lab/monocle3')

# Installing old version of SeuratWrappers to avoid conflict with monocle3

remotes::install_github('satijalab/seurat-wrappers@community-vignette')


```
# Introduction


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_EC_Clean <- readRDS(here("outputs", "Alports_Glom_EC.rds"))

DimPlot(SO_EC_Clean)

DimPlot(SO_EC_Clean, group.by = "group")

DimPlot(SO_EC_Clean, split.by = "group")

head(SO_EC_Clean@meta.data)

df<- FindAllMarkers(SO_EC_Clean, logfc.threshold = 0.25, only.pos = TRUE)


```
# Loading CDS object

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

seurat_obj <- SO_EC_Clean

DefaultAssay(seurat_obj) <- "RNA"

cds <- as.cell_data_set(seurat_obj)

cds <- cluster_cells(cds = cds, reduction_method = "UMAP")

# find all possible partitions
all_partitions <- unique(cds@clusters$UMAP$partitions)
all_partitions <- all_partitions[all_partitions != "1"]

# set all partitions to 1
cds@clusters$UMAP$partitions[cds@clusters$UMAP$partitions %in% all_partitions] <- "1"

# set all cells to one partition, otherwise it won't show the trajectory across partitions. Set use_partition to FALSE as below.
cds <- learn_graph(cds, use_partition = F)

# a helper function to identify the root principal points:
get_earliest_principal_node <- function(cds, time_bin="0"){
  cell_ids <- which(colData(cds)[, "seurat_clusters"] == time_bin)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}
cds <- order_cells(cds, root_pr_nodes=get_earliest_principal_node(cds))

cds <- order_cells(cds)

plot_cells(cds = cds,
           color_cells_by = "pseudotime",
           show_trajectory_graph = TRUE,
           label_groups_by_cluster=FALSE,
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           graph_label_size=3) 


cds_sub <- choose_graph_segments(cds)


```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_EC_Clean <- AddMetaData(
  object = SO_EC_Clean,
  metadata = cds_sub@0principal_graph_aux@listData$UMAP$pseudotime,
  col.name = "Pseudotime"
)

head(SO_EC_Clean@meta.data)

VlnPlot(SO_EC_Clean, features = "Pseudotime", group.by = "group")



```





## Mitochondrial dysfunction

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Mitochondria dysfunction`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Mitochondria_dysfunction")

VlnPlot(SO_EC_Clean, features = "Mitochondria_dysfunction1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Mitochondria_dysfunction1", group.by = "group", pt.size = 0)

kruskal.test(Mitochondria_dysfunction1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Mitochondria_dysfunction1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(mito_dysfunction = mean(Mitochondria_dysfunction1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df2 <- heatmap_data
```


## Mitochondrial division

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Mitochondrial Division Signaling Pathway`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Mitochondria_division")

VlnPlot(SO_EC_Clean, features = "Mitochondria_division1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Mitochondria_division1", group.by = "group", pt.size = 0)

kruskal.test(Mitochondria_division1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Mitochondria_division1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(mito_division = mean(Mitochondria_division1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df3 <- heatmap_data
```


## mTor

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`mTOR signaling`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "mTor")

VlnPlot(SO_EC_Clean, features = "mTor1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "mTor1", group.by = "group", pt.size = 0)

kruskal.test(mTor1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$mTor1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(mTor = mean(mTor1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df4 <- heatmap_data
```

## OxPhos

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Oxidative Phosphorylation`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "OxPhos")

VlnPlot(SO_EC_Clean, features = "OxPhos1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "OxPhos1", group.by = "group", pt.size = 0)

kruskal.test(OxPhos1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$OxPhos1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
  )

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(OxPhos = mean(OxPhos1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df5 <- heatmap_data
```

## PI Metabolism

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`PI Metabolism`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "PI_met")

VlnPlot(SO_EC_Clean, features = "PI_met1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "PI_met1", group.by = "group", pt.size = 0)

kruskal.test(PI_met1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$PI_met1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(PI_met = mean(PI_met1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df6 <- heatmap_data
```

## Mito protein degradation

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Mito protein degradation`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Mito_degradation")

VlnPlot(SO_EC_Clean, features = "Mito_degradation1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Mito_degradation1", group.by = "group", pt.size = 0)

kruskal.test(Mito_degradation1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Mito_degradation1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Mito_degradation = mean(Mito_degradation1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df7 <- heatmap_data
```

## eNoS signaling

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`eNOS signaling`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "eNos")

VlnPlot(SO_EC_Clean, features = "eNos1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "eNos1", group.by = "group", pt.size = 0)

kruskal.test(eNos1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$eNos1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(eNos = mean(eNos1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df8 <- heatmap_data

```

## Integration of energy metabolism

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Integration of energy metabolism`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Integration")

VlnPlot(SO_EC_Clean, features = "Integration1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Integration1", group.by = "group", pt.size = 0)

kruskal.test(Integration1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Integration1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Integration = mean(Integration1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df9 <- heatmap_data
```


## All Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# Convert all to data.frames (if not already)
df_list <- list(df2, df3, df4, df5, df6, df7, df8, df9)
df_list <- lapply(df_list, as.data.frame)

# Find common rownames across all dataframes
common_rows <- Reduce(intersect, lapply(df_list, rownames))

# Subset each dataframe to common rows, in the same order
df_list_sub <- lapply(df_list, function(df) df[common_rows, , drop = FALSE])

# Combine all dataframes by columns
merged_df <- do.call(cbind, df_list_sub)

# Convert to numeric matrix (important if some columns are factors/characters)
heatmap_data <- as.matrix(merged_df)

# Optional: check if any non-numeric columns exist and convert or remove them
if(any(!sapply(heatmap_data, is.numeric))) {
  stop("Non-numeric columns detected in heatmap_data. Please convert or remove them.")
}

# Plot heatmap
pheatmap(
  heatmap_data,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  scale = "column",  # scales each column (group) to mean=0, sd=1
  color = colorRampPalette(c("white", "red"))(50)
)


```

# Extract the data
scores <- SO_EC_Clean@meta.data$Integration1
groups <- SO_EC_Clean@meta.data$group

# Wilcoxon test (for two groups)
anova_result <- aov(SO_EC_Clean@meta.data$Integration1 ~ SO_EC_Clean@meta.data$group, data = SO_EC_Clean@meta.data)
summary(anova_result)
TukeyHSD(anova_result)


# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

