---
title: "365 Ambient RNA and Doublet Removal"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load_new_packages, echo=T, error=FALSE, message=FALSE, warning=FALSE}

if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("SoupX")) {install.packages("SoupX"); require("SoupX")}
if (!require("DoubletFinder")) {BiocManager::install('DoubletFinder'); require("DoubletFinder")}
if (!require("here")) {install.packages("here"); require("here")}
#if (!require("remotes")) {install.packages("remotes"); require("remotes")}

#remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')

options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```

# Analysis Notes

1) I'm not sure I can explain this, but there is some strange processing going on between the raw and filtered files and they don't have the same number of genes in them...so I would error out on the `sc = SoupChannel(filtered_matrix,toc)` code. To fix this, I filtered the tod to have the same genes as the toc with `rows_to_keep <- toc@Dimnames[[1]]` and `filtered_matrix <- tod[rows_to_keep, , drop = FALSE]`

# Dataset Pre-processing

## SoupX for Ambient RNA Removal
[Young and Behjati, 2020](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7763177/)

```{r pre_process, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sample <- "365"

# Load data and estimate soup profile
tod = Seurat::Read10X_h5(here("datasets", paste0(sample, "_sample_raw_feature_bc_matrix.h5"))) 
toc = Seurat::Read10X_h5(here("datasets", paste0(sample, "_sample_filtered_feature_bc_matrix.h5")))

rows_to_keep <- toc@Dimnames[[1]]
filtered_matrix <- tod[rows_to_keep, , drop = FALSE]

sc = SoupChannel(filtered_matrix,toc)

#Make the Seurat object from the filtered control data
SO <- Read10X_h5(here("datasets", paste0(sample, "_sample_filtered_feature_bc_matrix.h5")))
SO <- CreateSeuratObject(counts = SO, project = "Alports_EV")  
 
#Cluster the cells with Seurat
SO <- SCTransform(SO, verbose = F)
SO <- RunPCA(SO, verbose = F)
SO <- RunUMAP(SO, dims = 1:30, verbose = F)
SO <- FindNeighbors(SO, dims = 1:30, verbose = F)
SO <- FindClusters(SO, verbose = T)

meta <- SO@meta.data
umap <- SO@reductions$umap@cell.embeddings
clusters <- setNames(meta$seurat_clusters, rownames(meta))

#Sanity Check
length(clusters) #should be equal to nrow(sc$metaData)
nrow(sc$metaData)

sc <- setClusters(sc, clusters)
sc <- setDR(sc, umap)

#Estimate rho
sc = autoEstCont(sc)

#Clean the data
SO_out = adjustCounts(sc)
 
#Create a new Seurat Object out of the cleaned data
seurat.obj <- CreateSeuratObject(SO_out, min.cells = 3, min.features = 200)

```

## Pre-process for Doublet Finder

[(McGinnis et al., 2019)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6853612/)

```{r pre_process2, echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}

seurat.obj

seurat.obj[["percent.mt"]] <- PercentageFeatureSet(seurat.obj, pattern = "^mt-")

VlnPlot(seurat.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

VlnPlot(seurat.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)

# Minimal QC and Filtering (low quality cells) to let doublet find doublets
seurat.obj.f <- subset(seurat.obj,
                                  nFeature_RNA < 10000 &
                                  nCount_RNA < 30000 &
                                  percent.mt < 15 &
                                  nFeature_RNA > 1500
                                  ) 
seurat.obj.f

VlnPlot(seurat.obj.f, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

VlnPlot(seurat.obj.f, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0) & labs(subtitle = "ViolinPlot After Filtering")

# Pre-process standard workflow
seurat.obj.f <- NormalizeData(object = seurat.obj.f)
seurat.obj.f <- FindVariableFeatures(object = seurat.obj.f)
seurat.obj.f <- ScaleData(object = seurat.obj.f)
seurat.obj.f <- RunPCA(object = seurat.obj.f)
ElbowPlot(seurat.obj.f, ndims = 40)

# PCs between 15-20
seurat.obj.f <- FindNeighbors(object = seurat.obj.f, dims = 1:50)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 0.4)
seurat.obj.f <- RunUMAP(object = seurat.obj.f, dims = 1:50)
DimPlot(seurat.obj.f, reduction = "umap")

FeaturePlot(seurat.obj.f, "Nphs2") + ggtitle("Nphs2: Podocyte Marker")
FeaturePlot(seurat.obj.f, "Flt1") + ggtitle("Flt1: Endothelial Marker")
FeaturePlot(seurat.obj.f, "Pdgfrb") + ggtitle("Pdgfrb: Mesangial Marker")
FeaturePlot(seurat.obj.f, "Slc12a3") + ggtitle("Slc12a3: DCT Marker")
FeaturePlot(seurat.obj.f, "Ptprc") + ggtitle("Ptprc: Immune Marker")
FeaturePlot(seurat.obj.f, "Ncam1") + ggtitle("Ncam1: Parietal Epithelial Marker")

```

```{r find_parameters_1, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', results=F}

# Calculate each combination of pN and pK
sweep.res.list_seurat.obj.f <- paramSweep(seurat.obj.f, PCs = 1:20, sct = FALSE) 
```

```{r find_parameters_2, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

#Summarize each combination of pN and pK
sweep.stats_seurat.obj.f <- summarizeSweep(sweep.res.list_seurat.obj.f, GT = FALSE) 

#Select the pK that corresponds to max bcmvn to optimize doublet detection
bcmvn_seurat.obj.f <- find.pK(sweep.stats_seurat.obj.f)
pK <- bcmvn_seurat.obj.f %>% 
  filter(BCmetric == max(BCmetric)) %>%
  dplyr::select(pK) 

#See pK in the Values Environment
pK <- as.numeric(as.character(pK[[1]]))

```

 
```{r homotypic_doublets, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

 
# Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
annotations <- seurat.obj.f@meta.data$seurat_clusters  
 
homotypic.prop <- modelHomotypic(annotations)           
homotypic.prop
 
# 10X Multiplet Rate Table (the doublet ratio is # of cells recovered divided by 125000) https://kb.10xgenomics.com/hc/en-us/articles/360001378811-What-is-the-maximum-number-of-cells-that-can-be-profiled-
 
nrow(seurat.obj.f@meta.data)

nExp_poi <- round(nrow(seurat.obj.f@meta.data) # To calculate cell number
                  /125000              # To calculate the doublet ratio
                  *nrow(seurat.obj.f@meta.data)) #* 1.4
nExp_poi
nExp_poi_adj <- round(nExp_poi*(1-homotypic.prop))
 
```

## Doublet Finder

```{r doublet_finder, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
 
seurat.obj.f_doublets <- doubletFinder(seurat.obj.f,
                        PCs = 1:20,
                        pN = 0.25,
                        pK = pK,
                        nExp = nExp_poi_adj,
                        reuse.pANN = FALSE, sct = FALSE)

head(seurat.obj.f_doublets@meta.data)

colnames(seurat.obj.f_doublets@meta.data)[7] <- "pANN"
colnames(seurat.obj.f_doublets@meta.data)[8] <- "DF.class"
head(seurat.obj.f_doublets@meta.data)
table(seurat.obj.f_doublets@meta.data$DF.class)

DimPlot(seurat.obj.f_doublets, group.by = "DF.class")

VlnPlot(seurat.obj.f_doublets, "nFeature_RNA", group.by = "DF.class")

VlnPlot(seurat.obj.f_doublets, "nCount_RNA", group.by = "DF.class")

```


### Subset singlets

```{r subset, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

seurat.obj.f_singlets <- subset(seurat.obj.f_doublets, DF.class == "Singlet")
seurat.obj.f_singlets
DimPlot(seurat.obj.f_singlets, reduction = "umap", label = TRUE)

```

## Add Meta data

```{r meta data, echo=TRUE, error=FALSE, fig.align= 'center', message=FALSE, warning=FALSE}
# Create Seurat object and add metadata

seurat.obj.f_singlets <- AddMetaData(object = seurat.obj.f_singlets, metadata = sample, col.name = "sample") #Change
#seurat.obj.f_singlets[["percent.mt"]] <- PercentageFeatureSet(seurat.obj.f_singlets, pattern = "^mt-")

head(seurat.obj.f_singlets@meta.data)

```


# Initial Cluster Identification

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center', fig.height=10, fig.width=8}


markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Cdh13",        # dTL1
                      "Slc14a2",      # dTL2
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs2",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b",        # Uro
                      "Top2a"         # Proliferation
)
                      
DotPlot(seurat.obj.f_singlets,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5)+
coord_flip()

```

## Intermediate Output Files

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(seurat.obj.f_singlets, here("outputs", paste0("Alports_", sample, ".rds")))

```

# Session Info

```{r  echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, results=T}

sessionInfo()

```


