---
title: "Alports EC: Geneset Quantification"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}

# library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)
library(pheatmap)


options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()

```
# Introduction


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_EC_Clean <- readRDS(here("outputs", "Alports_Glom_EC.rds"))

DimPlot(SO_EC_Clean)

```
# Identifying Reversal Genes

## Set the conditions

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

wb <- loadWorkbook(here("datasets", "IPA pathways _updated_08_18_2025 .xlsx"))

# Get sheet names
sheets <- getSheetNames(here("datasets", "IPA pathways _updated_08_18_2025 .xlsx"))

# Initialize an empty list
pathway_list <- list()

# Loop through each sheet
for (sheet in sheets) {
  # Read the entire sheet
  data <- read.xlsx(wb, sheet = sheet, colNames = FALSE)
  
  # Extract the name from cell A1
  list_name <- as.character(data[1, 1])
  
  # Extract all values from column B, starting from row 2 (since row 1 is A1 and likely the header of column B)
  values <- data[-1, 2, drop = TRUE]
  
  # Remove NAs and convert to character vector
  values <- as.character(values[!is.na(values)])
  
  # Add to list using the name from A1
  pathway_list[[list_name]] <- values
}

# View the list
str(pathway_list)

pathway_list
```

## Mitochondrial dysfunction

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Mitochondria dysfunction`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Mitochondria_dysfunction")

VlnPlot(SO_EC_Clean, features = "Mitochondria_dysfunction1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Mitochondria_dysfunction1", group.by = "group", pt.size = 0)

kruskal.test(Mitochondria_dysfunction1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Mitochondria_dysfunction1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(mito_dysfunction = mean(Mitochondria_dysfunction1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df2 <- heatmap_data
```


## Mitochondrial division

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Mitochondrial Division Signaling Pathway`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Mitochondria_division")

VlnPlot(SO_EC_Clean, features = "Mitochondria_division1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Mitochondria_division1", group.by = "group", pt.size = 0)

kruskal.test(Mitochondria_division1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Mitochondria_division1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(mito_division = mean(Mitochondria_division1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df3 <- heatmap_data
```


## mTor

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`m`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "mTor")

VlnPlot(SO_EC_Clean, features = "mTor1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "mTor1", group.by = "group", pt.size = 0)

kruskal.test(mTor1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$mTor1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(mTor = mean(mTor1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df4 <- heatmap_data
```

## OxPhos

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Oxidative Phosphorylation`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "OxPhos")

VlnPlot(SO_EC_Clean, features = "OxPhos1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "OxPhos1", group.by = "group", pt.size = 0)

kruskal.test(OxPhos1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$OxPhos1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
  )

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(OxPhos = mean(OxPhos1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df5 <- heatmap_data
```

## PI Metabolism

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`PI Metabolism`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "PI_met")

VlnPlot(SO_EC_Clean, features = "PI_met1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "PI_met1", group.by = "group", pt.size = 0)

kruskal.test(PI_met1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$PI_met1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(PI_met = mean(PI_met1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df6 <- heatmap_data
```

## Mito protein degradation

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Mito protein degradation`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Mito_degradation")

VlnPlot(SO_EC_Clean, features = "Mito_degradation1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Mito_degradation1", group.by = "group", pt.size = 0)

kruskal.test(Mito_degradation1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Mito_degradation1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)

group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Mito_degradation = mean(Mito_degradation1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df7 <- heatmap_data
```

## eNoS signaling

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`eNOS signaling`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "eNos")

VlnPlot(SO_EC_Clean, features = "eNos1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "eNos1", group.by = "group", pt.size = 0)

kruskal.test(eNos1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$eNos1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(eNos = mean(eNos1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df8 <- heatmap_data

```

## Integration of energy metabolism

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Integration of energy metabolism`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Integration")

VlnPlot(SO_EC_Clean, features = "Integration1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Integration1", group.by = "group", pt.size = 0)

kruskal.test(Integration1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Integration1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Integration = mean(Integration1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df9 <- heatmap_data
```

## Accumulation of Lipids

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Accumulation of Lipids`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Lipids")

VlnPlot(SO_EC_Clean, features = "Lipids1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Lipids1", group.by = "group", pt.size = 0)

kruskal.test(Lipids1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Lipids1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Lipids = mean(Lipids1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df10 <- heatmap_data
```

##  Metabolism of Membrane Lipid Derivatives

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Metabolism of Membrane Lipid Derivatives`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Derivatives")

VlnPlot(SO_EC_Clean, features = "Derivatives1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Derivatives1", group.by = "group", pt.size = 0)

kruskal.test(Derivatives1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Derivatives1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Derivatives = mean(Derivatives1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df11 <- heatmap_data
```

## Synthesis of Lipid

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`Synthesis of Lipid`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "Synthesis")

VlnPlot(SO_EC_Clean, features = "Synthesis1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "Synthesis1", group.by = "group", pt.size = 0)

kruskal.test(Synthesis1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$Synthesis1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(Synthesis = mean(Synthesis1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df12 <- heatmap_data
```

## ER-Mitochonrial Communication Signaling 

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`ER-Mitochonrial Communication Signaling `)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "ER_Mito")

VlnPlot(SO_EC_Clean, features = "ER_Mito1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "ER_Mito1", group.by = "group", pt.size = 0)

kruskal.test(ER_Mito1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$ER_Mito1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(ER_Mito = mean(ER_Mito1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df13 <- heatmap_data
```

## NAD signaling 

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1 <- list(pathway_list$`NAD signaling`)

SO_EC_Clean <- AddModuleScore(object = SO_EC_Clean , features = df1, name = "NAD")

VlnPlot(SO_EC_Clean, features = "NAD1", group.by = "group")

VlnPlot(SO_EC_Clean, features = "NAD1", group.by = "group", pt.size = 0)

kruskal.test(NAD1 ~ group, data = SO_EC_Clean@meta.data)

pairwise.wilcox.test(
  x = SO_EC_Clean@meta.data$NAD1,
  g = SO_EC_Clean@meta.data$group,
  p.adjust.method = "BH"  # or "bonferroni", "holm", etc.
)


group_means <- SO_EC_Clean@meta.data %>%
  group_by(group) %>%
  summarize(NAD = mean(NAD1, na.rm = TRUE))

heatmap_data <- as.matrix(group_means %>% column_to_rownames("group"))

pheatmap(heatmap_data, cluster_rows = FALSE, cluster_cols = FALSE, scale = "column",
         color = colorRampPalette(c("white", "red"))(50))

df14 <- heatmap_data
```


## All Pathways

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# Convert all to data.frames (if not already)
df_list <- list(df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12, df13, df14)
df_list <- lapply(df_list, as.data.frame)

# Find common rownames across all dataframes
common_rows <- Reduce(intersect, lapply(df_list, rownames))

# Subset each dataframe to common rows, in the same order
df_list_sub <- lapply(df_list, function(df) df[common_rows, , drop = FALSE])

# Combine all dataframes by columns
merged_df <- do.call(cbind, df_list_sub)

# Convert to numeric matrix (important if some columns are factors/characters)
heatmap_data <- as.matrix(merged_df)

# Optional: check if any non-numeric columns exist and convert or remove them
if(any(!sapply(heatmap_data, is.numeric))) {
  stop("Non-numeric columns detected in heatmap_data. Please convert or remove them.")
}

# Plot heatmap
pheatmap(
  heatmap_data,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  scale = "column",  # scales each column (group) to mean=0, sd=1
  color = colorRampPalette(c("white", "red"))(50)
)


```

## All Pathways
### Remove the 6M Group

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

# Convert all to data.frames (if not already)
df_list <- list(df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12, df13, df14)
df_list <- lapply(df_list, as.data.frame)

# Find common rownames across all dataframes
common_rows <- Reduce(intersect, lapply(df_list, rownames))

# Subset each dataframe to common rows, in the same order
df_list_sub <- lapply(df_list, function(df) df[common_rows, , drop = FALSE])

# Combine all dataframes by columns
merged_df <- do.call(cbind, df_list_sub)

# Convert to numeric matrix (important if some columns are factors/characters)
heatmap_data <- as.matrix(merged_df)

heatmap_data <- heatmap_data[rownames(heatmap_data) != "6M_Alp", ]

# Optional: check if any non-numeric columns exist and convert or remove them
if(any(!sapply(heatmap_data, is.numeric))) {
  stop("Non-numeric columns detected in heatmap_data. Please convert or remove them.")
}

# Plot heatmap
pheatmap(
  heatmap_data,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  scale = "column",  # scales each column (group) to mean=0, sd=1
  color = colorRampPalette(c("white", "red"))(50)
)


```



# Extract the data
scores <- SO_EC_Clean@meta.data$Integration1
groups <- SO_EC_Clean@meta.data$group

# Wilcoxon test (for two groups)
anova_result <- aov(SO_EC_Clean@meta.data$Integration1 ~ SO_EC_Clean@meta.data$group, data = SO_EC_Clean@meta.data)
summary(anova_result)
TukeyHSD(anova_result)


# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

