---
title: "Alports EC: Quantify EC Reversal Genes"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("clusterProfiler")) {BiocManager::install('clusterProfiler'); require("clusterProfiler")}
if (!require("enrichplot")) {BiocManager::install('enrichplot'); require("enrichplot")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for data frame transformation
if (!require("gplots")) {install.packages("gplots"); require("gplots")}
if (!require("ggvenn")) {install.packages("ggvenn"); require("ggvenn")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}

library("EnhancedVolcano")
library(UpSetR)
library(ComplexHeatmap)
#install.packages("ggtext")
library(ggtext)



options(future.globals.maxSize = 74 * 1024^3) # 55 GB
getOption("future.globals.maxSize") #59055800320

here()
```
# Introduction


```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

SO_EC_Clean <- readRDS(here("outputs", "Alports_Glom_EC.rds"))

DimPlot(SO_EC_Clean)

```
# Identifying Reversal Genes

## Set the conditions

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Condition_Baseline <- "WT"

Condition_A <- "4M_Alp"
  
Condition_B <- "4M_Alp_1EV"

Condition_C <- "4M_Alp_2EV"

Condition_D <- "6M_Alp"

```

## Create DEG lists

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

Idents(SO_EC_Clean) <- SO_EC_Clean@meta.data$subclass_EC

df1 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_A, ident.2 = Condition_Baseline, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df2 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_B, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df3 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_C, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)

df4 <- FindMarkers(SO_EC_Clean, ident.1 = Condition_D, ident.2 = Condition_A, group.by = "group", subset.ident = "EC_glom", logfc.threshold = 0.1, min.pct = 0.1)



```

## Split by Up and Down

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1_pos <- df1 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df1_pos

df1_neg <- df1 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df1_neg

df2_pos <- df2 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df2_pos

df2_neg <- df2 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df2_neg

df3_pos <- df3 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df3_pos

df3_neg <- df3 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df3_neg

df4_pos <- df4 %>% filter(avg_log2FC > 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df4_pos

df4_neg <- df4 %>% filter(avg_log2FC < 0) %>% filter(p_val_adj < 0.05) %>% arrange(desc(abs(avg_log2FC)))

df4_neg

```

## Reversal Venn Diagrams

### Up -> Down

#### 4M vs 4M_1EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_pos)
set2 <- rownames(df2_neg)
venn <- ggvenn(list('Up in 4M' = set1, 'Down in 1EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_1EV <- test$'A:B'

```

#### 4M vs 4M_2EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_pos)
set2 <- rownames(df3_neg)
venn <- ggvenn(list('Up in 4M' = set1, 'Down in 2EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_2EV <- test$'A:B'

```

#### 4M vs 6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_pos)
set2 <- rownames(df4_neg)
venn <- ggvenn(list('Up in 4M' = set1, 'Down in 6M' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

genelist_6M_Alp <- test$'A:B'

```

# Upset Plot Down in 4M vs Up in Others

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1_pos <- rownames(df1_pos)
df2_neg <- rownames(df2_neg)
df3_neg <- rownames(df3_neg)
df4_neg <- rownames(df4_neg)

my_lists <- list("4M_Alp" = df1_pos, "4M_Alp_1EV" = df2_neg, "4M_Alp_2EV" = df3_neg , "6M_Alp" = df4_neg)

str(my_lists)

list_to_matrix(my_lists)

dfX <- list_to_matrix(my_lists)

m1 <-  make_comb_mat(my_lists)
m1

UpSet(m1, comb_order = order(comb_size(m1)))

```

## Sanity Check

### Check for genes with reversal in all treatments

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df_sanity <- extract_comb(m1,"1111")

df_sanity

df_sanity2 <- head(df_sanity, 10)

for (i in df_sanity2) {
  A <- VlnPlot(SO_EC_Clean, i, split.by = "group")
  print(A)
}

```




### Down -> Up

#### 4M vs 4M_1EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_neg)
set2 <- rownames(df2_pos)
venn <- ggvenn(list('Down in 4M' = set1, 'Up in 1EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_1EV <- test$'A:B'

```

#### 4M vs 4M_2EV

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_neg)
set2 <- rownames(df3_pos)
venn <- ggvenn(list('Down in 4M' = set1, 'Up in 2EV' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

genelist_4M_Alp_2EV <- test$'A:B'

```

#### 4M vs 6M

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

set1 <- rownames(df1_neg)
set2 <- rownames(df4_pos)
venn <- ggvenn(list('Down in 4M' = set1, 'Up in 6M' = set2))
print(venn)

lst <- list(set1, set2)
ItemsList <- venn(lst, show.plot = TRUE)

test <- attributes(ItemsList)$intersections

genelist_6M_Alp <- test$'A:B'

```

# Upset Plot Up in 4M vs Down in Others

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df1_neg <- rownames(df1_neg)
df2_pos <- rownames(df2_pos)
df3_pos <- rownames(df3_pos)
df4_pos <- rownames(df4_pos)

my_lists <- list("4M_Alp" = df1_neg, "4M_Alp_1EV" = df2_pos, "4M_Alp_2EV" = df3_pos , "6M_Alp" = df4_pos)

str(my_lists)

list_to_matrix(my_lists)

dfX <- list_to_matrix(my_lists)

m1 <-  make_comb_mat(my_lists)
m1

UpSet(m1, comb_order = order(comb_size(m1)))

```

## Sanity Check

### Check for genes with reversal in all treatments

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

df_sanity <- extract_comb(m1,"1111")

df_sanity

df_sanity2 <- head(df_sanity, 10)

for (i in df_sanity2) {
  A <- VlnPlot(SO_EC_Clean, i, split.by = "group")
  print(A)
}

```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

